### 2024-05-13 index.qmd
init

---
title: "Current Sheet Theory"
code-links:
#   - repo
  - text: Tana
    href: https://app.tana.inc?nodeid=OzzX0sVA4_6X
  - text: Zotero
    href: zotero://select/library/collections/DZHNUFNV
---

## References

- @steinhauerMultifluidModelOnedimensional2008

### 2024-05-14 index.qmd
update

@@
 code-links:
-#   - repo
+  - repo
   - text: Tana
@@
 
+One-dimensional geometry in which all variables depend only on the z (normal) coordinate.
+
+$$
+d B_y / d z=\sum_\alpha(-n_\alpha u_{\alpha x}+\Gamma_\alpha B_x / B_z)
+$$
+
+$$
+d B_x / d z=\sum_\alpha(n_\alpha u_{\alpha y}-\Gamma_\alpha B_y / B_z)
+$$
+
+$$
+\Gamma_e=\sum_\alpha n_\alpha u_{\alpha z}=\sum_\alpha \Gamma_\alpha
+$$
+
+We are interested in solution that have $d B_y / d z \neq 0$
+
 ## References

### 2024-05-16 main.jl
init update scripts to compute

@@ -1,0 +1,29 @@
+using BoundaryValueDiffEq
+using DifferentialEquations
+
+function system!(du, u, p, t)
+    n, Bx, By, ux, uy = u
+    dBy = -n * ux
+    dBx = n * uy
+    dux = By * (n - 1)
+    duy = Bx * (1 - n)
+    dn = n * (ux * By - uy * Bx) / (- Bz^2/n^2 + betaI * gamma * n^(gamma - 1))
+    du .= [dn, dBx, dBy, dux, duy]
+end
+
+function bc2a!(resid_a, u, p, t)
+    resid_a[2] = u[2] 
+end
+
+function bc2b!(resid_b, u, p, t)
+    resid_b[1] = u[1]
+    resid_b[2] = u[2] - 1
+    resid_b[3] = u[3]
+    resid_b[4] = u[4]
+    resid_b[5] = u[5]
+end
+
+span = (0.0, 10.0)
+u0 = [1.5, 0.0, 1, 0.0, 1]
+
+bvp = TwoPointBVProblem(system!, (bc2a!, bc2b!), u0, span;)

### 2024-05-24 index.qmd
new: complete derivation and figures

--- 2024-05-13 index.qmd
@@
 
+:::{.hide}
+\newcommand{\vect}[1]{\mathbf{#1}}
+:::
+
+## Plasma-Field Equations for a One-Dimensional Multifluid
+
+Consider a collisionless, steady state plasma composed of multiple fluid groups, and follow the notation in @steinhauerMultifluidModelOnedimensional2008, we have $3N$ equations for $N$ ion species:
+
+$$
+\begin{aligned}
+Γ_α \frac{d u_{α x}}{d z} &= n_α u_{α y} B_z - Γ_α B_y
+\\
+Γ_α \frac{d u_{α y}}{d z} &= Γ_α B_x-n_α u_{α x} B_z
+\\
+Γ_α \frac{d u_{α z}}{d z} &= -\frac{1}{2} \frac{d p_α}{d z}-n_α \frac{d \phi}{d z}+n_α u_{α x} B_y-n_α u_{α y} B_x
+\end{aligned}
+$${#eq-momentum}
+
+Ampere’s law connects the fields and the flow components and note that electron motion is along the field lines $\vect{u_e} = Γ_e \vect{B} / (n_e B_z)$ :
+
+$$
+d B_y / d z = - J_x = - \sum_α n_α u_{α x} + n_e u_{e x} = -n u_x+\Gamma_e B_x / B_z
+$${#eq-Jx}
+
+$$
+d B_x / d z = J_y = \sum_α n_α u_{α y} - n_e u_{e y} = n u_y-\Gamma_e B_y / B_z
+$${#eq-Jy}
+
+This is a system of $3N+2$ equations for $4N+2$ unknown dependent variables: $B_x$, $B_y$, and $N$ each of $n$, $p$ $u_x$, and $u_y$.
+
+- $B_z, Γ_α$ are constant.
+
+- Combine momentum equation (x) (@eq-momentum) and Ampere's law (y) (@eq-Jy) with condition that the constant of integration vanishes at the current sheet center yield
+
+  $$
+  B_x B_z = \sum n_α u_{α,x} u_{α,z} = \sum Γ_α u_{α,x}
+  $$
+
+  - With asymptotic condition that the derivative goes zero, from momentum equation we have
+
+    $$
+    {B_z}^2 = \sum \Gamma_{\alpha}^2/n_{\alpha}(\infty)
+    $${#eq-asym1}
+
+In the following, we will consider a system with two ion species.
+
+## Force-free current sheet
+
+We are interested in solutions with $B_x^2 + B_y^2 = B_0^2 = const$ (force-free current sheet). From Ampere's law (@eq-Jx) and (@eq-Jy), we have:
+
+$$
+n (u_x B_y  - u_y B_x) = 0
+$$
+
+And $n_1 u_{1x} B_y = n_1 u_{1y} B_x + C_1$ and $n_2 u_{2x} B_y = n_2 u_{2y} B_x  - C_1$. Assuming $C_1 = 0$, from first two equations in (@eq-momentum), we have:
+
+$$
+u_{αx}^2 + u_{αy}^2 = const
+$$
+
+::: {#def-1}
+Express the quantities of the second ion species relative to the first species
+
+$$
+u_{2x} := λ_x u_{1x}, \quad u_{2y} := λ_y u_{1y}, \quad u_{2z} := λ_z u_{1z}, \quad n_2 := λ_n n_1
+$$
+
+$$
+Λ_x := 1 + λ_x λ_n, \quad Λ_y := 1 + λ_y λ_n, \quad Λ_z := 1 + λ_z λ_n
+$$
+
+Notes
+
+- $Γ_e = (1 + λ_z λ_n) Λ_1 =  Λ_z  Γ_1$. 
+- $Λ_z$ is constant while $Λ_x, Λ_y$ are not.
+
+:::
+
+::: {#def-2}
+Force-free condition let us express the magnetic field and velocity in terms of the angle $θ$:
+
+$$
+B_x = B_0 \cos θ, \quad B_y = B_0 \sin θ
+$$
+
+$$
+u_{1x} = u_{1} \cos θ_1, \quad u_{1y} = u_{1} \sin θ_1
+$$
+:::
+
+$C_1=0$ immediately implies $θ = θ_1 + k \pi$ and $Λ_x=Λ_y$.
+
+The momentum equations (@eq-momentum) become:
+
+$$
+- u_1 \sin θ_1 θ_1' = n_1 u_1 \sin θ_1 B_z / Γ_1 - B_0 \sin θ
+$$$$
+\Rightarrow 
+θ' = - \frac{n_1 B_z}{Γ_1} \pm \frac{B_0}{u_1}
+$${#eq-theta1}
+
+- Note that $θ_1, n_1$ are dependent variables, and $u_1, B_0, B_z, Γ_1$ are constants determined by the system.
+- **Given $n_1$, we could solve $θ_1$ from the above equation.**
+- The derivative of $θ$ goes to zero at infinity, gives us a relation between $u_1$ and $B_0$:
+
+  $$
+  u_1 = \pm \frac{B_0 Γ_1}{B_z n_1}
+  $$
+
+The Ampere's law (@eq-Jx) become:
+
+$$
+B_0 \sin θ θ' = Λ_y n_1 u_1 \sin θ_1 - Λ_z Γ_1 B_0 \sin θ / B_z
+$$$$
+\Rightarrow 
+θ' = \mp \frac{Λ_y n_1 u_1 }{B_0} + \frac{Λ_z Γ_1}{B_z}
+$${#eq-theta}
+
+By equating the above two equations, we could get a relation between $n_1$ and $Λ_y$:
+
+$$
+Λ_y = Λ_y(n_1) = -\frac{B_0 \left(-n_1 u_1 B_z^2+B_0 \Gamma _1 B_z-\Gamma _1^2 u_1 \Lambda _z\right)}{\Gamma _1 n_1 u_1^2 B_z}
+$$
+
+### Solutions
+
+Normalize the density by $n_1(z=\infty) = 1$ and the magnetic field by $B_z = 1$.
+
+Assuming density profile $n_1(z) \to \frac{c}{\left(\frac{z}{\delta }\right)^2+1}+1$
+
+We have
+
+$$
+\begin{aligned}
+\varphi (z) &\to \frac{\pi  \Gamma _1-2 c \delta  \tan ^{-1}\left(\frac{z}{\delta }\right)}{2 \Gamma _1}
+\\
+B_x &\to B_0 \cos \left(\frac{\pi  \Gamma _1-2 c \delta  \tan ^{-1}\left(\frac{z}{\delta }\right)}{2 \Gamma _1}\right)
+\end{aligned}
+$$
+
+For the simplest case $λ_n = 1, λ_z = - 1$, from @eq-asym1, we have $Γ_1 = - B_z / \sqrt{2}$.
+
+We could also normalize the system length by $\delta = 1$. So now the system could be fully determined by $c, B_0$.
+
+Given $c = 1/\sqrt{2}$, we have $B_y \to 0$ as $z \to \infty$. Profiles are plotted below.
+
+![](figures/profiles.jpg)
+
+![](figures/J_profiles.jpg)
+
+The multi-fluid effect is clearly demonstrated by $Λ_y - 1$.
+
+```mathematica
+Manipulate[
+ Block[{B0 = B0i, c = ci},
+  p1 = Plot[
+    Evaluate[{n[z], Bx[z], By[z], u1x[z], u1y[z]} /. sols //. 
+      rulesn],
+    {z, -zmax, zmax},
+    PlotLegends -> {n, Subscript[B, x], Subscript[B, y], Subscript[u, 
+      x], Subscript[u, y]}];
+  Export["figures/profiles.jpg", p1];
+  ],
+
+ {{ci, 1/Sqrt[2]}, 0, 2},
+ {{B0i, 2}, 0.5, 2},
+ {{zmax, 10}, 5, 30}
+ ]
+
+Manipulate[
+ Block[{B0 = B0i, c = ci, Λz = Λz0},
+  p = Plot[
+    Evaluate[{Λy - 1, Jxi, Jx, Jyi, Jy} /. sols //. 
+      rulesn],
+    {z, -zmax, zmax},
+    PlotLegends -> {Subscript[Λ, y] - 1, Subscript[J, 
+      x, i], Subscript[J, x], Subscript[J, y, i], Subscript[J, y]}, 
+    PlotRange -> All];
+  Export["figures/J_profiles.jpg", p];
+  p
+  ],
+
+ {{ci, 1/Sqrt[2]}, 0, 2},
+ {{B0i, 2}, 0.5, 2},
+ {Λz0, 0, 1},
+ {{zmax, 10}, 5, 30}
+]
+```
+
+## Questions
+
+- Why not motionless electrons?
+
 ## References

### 2024-05-24 index.qmd
update
chore: refactor

@@
 code-links:
-  - repo
   - text: Tana
@@
 
-We are interested in solutions with $B_x^2 + B_y^2 = B_0^2 = const$ (force-free current sheet). From Ampere's law (@eq-Jx) and (@eq-Jy), we have:
+We are interested in solutions with $B_x^2 + B_y^2 = B_0^2 = const$ (force-free current sheet).
+This provides another equation for the system, now we have 3*2+2+1=9 equations for 10 unknowns (the system is close to be fully determined).
+From Ampere's law (@eq-Jx) and (@eq-Jy), we have:
 
@@
 
-$C_1=0$ immediately implies $θ = θ_1 + k \pi$ and $Λ_x=Λ_y$.
+$C_1=0$ immediately implies $θ = θ_1 + k \pi$ and $Λ_x=Λ_y$. #TODO: Prove $C_1=0$ is like a boundary condition not a equation.
 
@@
   $$
-  u_1 = \pm \frac{B_0 Γ_1}{B_z n_1}
-  $$
+  u_1 = \pm \frac{B_0 Γ_1}{B_z n_1(\infty)}
+  $${#eq-u1}
 
@@
 
+The system $S$ could be fully determined by $λ_n, λ_z, c, B_0$ parameters, provided $n(z)$ profiles.
+
+$$S=S(z; n(z); λ_n, λ_z, c, B_0)$$
+
 Assuming density profile $n_1(z) \to \frac{c}{\left(\frac{z}{\delta }\right)^2+1}+1$
@@
 
-Given $c = 1/\sqrt{2}$, we have $B_y \to 0$ as $z \to \infty$. Profiles are plotted below.
+Given $c = 1/\sqrt{2}$, we have $B_y \to 0$ as $z \to \infty$. Profiles are plotted below for $B_0 = 2$.
 
@@
   p1 = Plot[
-    Evaluate[{n[z], Bx[z], By[z], u1x[z], u1y[z]} /. sols //. 
-      rulesn],
+    Evaluate[{n[z], Bx[z], By[z], u1x[z], u1y[z]} /. sols //. rulesn],
     {z, -zmax, zmax},
-    PlotLegends -> {n, Subscript[B, x], Subscript[B, y], Subscript[u, 
-      x], Subscript[u, y]}];
+    PlotLegends -> {n, Subscript[B, x], Subscript[B, y], Subscript[u, x], Subscript[u, y]}
+  ];
   Export["figures/profiles.jpg", p1];
+  p1
   ],
@@
 
-## Questions
-
-- Why not motionless electrons?
-
 ## References

### 2024-06-03 index.qmd
fix typo
chores
update motivations

@@
 title: "Current Sheet Theory"
+format:
+  html: default
+  pptx:
+    shift-heading-level-by: -1
 code-links:
@@
 :::
+
+## Motivation
+
+In MHD theory, for rotational rotational discontinuity, we have this following relationship
+
+$$
+[[\vec{U}]]=\left(\frac{\xi \rho}{\mu_0}\right)^{1 / 2}\left[\left[\frac{\vec{B}}{\rho}\right]\right]
+$$
+
+where $\xi \equiv 1-\frac{p_{\parallel}-p_{\perp}}{B^2 / \mu_0}$.
+
+However, observations of solar wind discontinuities reveals discrepancies between
+
+- Alfven velocity and plasma velocity change across discontinuities
+- Anisotropic MHD theory-predicted and directly measured ion anisotropies
+
+Multifluid theory is proposed to address these discrepancies. The basic idea is that we could have zero bulk velocity with non zero pressure.
+
+### Previous Work
+
+On the extreme of multiple fluid theory, we have the kinetic Harris model for the current sheet.
+
+> Consider several deficiencies of the Harris model: (1) there is no normal field, Bz = 0, so that no particles cross the sheet from asymptotic sources; (2) the unrealistic assumption of uniform cross-tail drift velocity, uy; (3) vanishing particle density far from the sheet; (4) assumption of a simple Maxwellian velocity distribution; (5) further, one study showed that atypical current sheet structures with bifurcated current density occur 17% of the time, and even for typical center-peaked sheets, strongly peaked structures frequently appear in the center, which also differs from the Harris profile [Asano et al., 2005].
+
+@steinhauerMultifluidModelOnedimensional2008 built a multifluid model to study the steady state magnetotail current sheet.
+
+![Ion trajectories in model magnetic field.](figures/jgra18999-fig-0001.png)
+
+For the symmetric three-fluid system with $B_y=0$,
+
+![Magnetic field, current density, ion density and out-of-plane flow for Bz = 0.1, βe = 0.5, βi = 1.
+Solid lines are fluid model results and dot symbols are from the Harris sheet solution.](figures/jgra18999-fig-0003.png)
+
+Here, we are interested in the case of a force-free current sheet and start from the simplest case of two ion species.
 
@@
 $$
-u_{2x} := λ_x u_{1x}, \quad u_{2y} := λ_y u_{1y}, \quad u_{2z} := λ_z u_{1z}, \quad n_2 := λ_n n_1
+u_{αx} := λ_{αx} u_{1x}, \quad u_{αy} := λ_{αy} u_{1y}, \quad u_{αz} := λ_{αz} u_{1z}, \quad n_α := λ_{αn} n_1
 $$
@@
 $$
-Λ_x := 1 + λ_x λ_n, \quad Λ_y := 1 + λ_y λ_n, \quad Λ_z := 1 + λ_z λ_n
+Λ_x := 1 + \sum λ_{αx} λ_{αn}, \quad Λ_y := 1 + \sum λ_{αy} λ_{αn}, \quad Λ_z := \sum 1 + λ_{αz} λ_{αn}
 $$
@@
 
-The Ampere's law (@eq-Jx) become:
+The Ampere's law (@eq-Jy) become:
 
 $$
-B_0 \sin θ θ' = Λ_y n_1 u_1 \sin θ_1 - Λ_z Γ_1 B_0 \sin θ / B_z
-$$$$
+- B_0 \sin θ θ' = Λ_y n_1 u_1 \sin θ_1 - Λ_z Γ_1 B_0 \sin θ / B_z
+$$
+
+$$
 \Rightarrow 
@@
 
-The system $S$ could be fully determined by $λ_n, λ_z, c, B_0$ parameters, provided $n(z)$ profiles.
+The system $S$ could be fully determined by $λ_n, λ_z, B_0$ parameters, provided $n(z)$ profiles.
 
-$$S=S(z; n(z); λ_n, λ_z, c, B_0)$$
+$$
+S=S(z; n(z); λ_n, λ_z, B_0)
+$$
 

### 2024-06-03 index.qmd
grpMeeting

@@
 ---
-title: "Current Sheet Theory"
+title: "Multifluid model of a one-dimensional steady state force-free current sheet"
 format:
@@
   pptx:
-    shift-heading-level-by: -1
+    slide-level: 3
+    incremental: true
 code-links:
@@
 
-## Motivation
+## Introduction
+
+### Motivation
 
@@
 where $\xi \equiv 1-\frac{p_{\parallel}-p_{\perp}}{B^2 / \mu_0}$.
+
+---
 
@@
 
-> Consider several deficiencies of the Harris model: (1) there is no normal field, Bz = 0, so that no particles cross the sheet from asymptotic sources; (2) the unrealistic assumption of uniform cross-tail drift velocity, uy; (3) vanishing particle density far from the sheet; (4) assumption of a simple Maxwellian velocity distribution; (5) further, one study showed that atypical current sheet structures with bifurcated current density occur 17% of the time, and even for typical center-peaked sheets, strongly peaked structures frequently appear in the center, which also differs from the Harris profile [Asano et al., 2005].
+- no normal field, $B_z = 0$
+- uniform cross-tail drift velocity, $u_y$
+- vanishing particle density far from the sheet
+
+<!-- > Consider several deficiencies of the Harris model: (1) there is no normal field, Bz = 0, so that no particles cross the sheet from asymptotic sources; (2) the unrealistic assumption of uniform cross-tail drift velocity, uy; (3) vanishing particle density far from the sheet; (4) assumption of a simple Maxwellian velocity distribution; (5) further, one study showed that atypical current sheet structures with bifurcated current density occur 17% of the time, and even for typical center-peaked sheets, strongly peaked structures frequently appear in the center, which also differs from the Harris profile [Asano et al., 2005].
+-->
 
@@
 ![Ion trajectories in model magnetic field.](figures/jgra18999-fig-0001.png)
+
+---
 
@@
 $${#eq-Jy}
+
+---
 
@@
 
+---
+
 ::: {#def-1}
@@
 :::
+
+---
 

### 2024-06-07 index.qmd
new: add velocity profiles for identical asympotic magnetic field

@@
 Multifluid theory is proposed to address these discrepancies. The basic idea is that we could have zero bulk velocity with non zero pressure.
+
+#### MHD in a nutshell
+
+$$
+\begin{aligned}
+& \frac{\partial \rho}{\partial t}+(\vec{V} \cdot \nabla) \rho+\rho \nabla \cdot \vec{V}=0 \\
+& \rho\left[\frac{\partial \vec{V}}{\partial t}+(\vec{V} \cdot \nabla) \vec{V}\right]+\nabla p=\vec{J} \times \vec{B} \\
+& p=\alpha \rho^\gamma \\
+& \nabla \times \vec{B}=\mu_0 \vec{J} \\
+& \frac{\partial \vec{B}}{\partial t}=\nabla \times(\vec{V} \times \vec{B})
+\end{aligned}
+$$
 
@@
 
+For the same asymptotic magnetic field, it is interesting to see how the plasma profiles change with different system parameters.
+
+$$
+\text{cond}(n(z), λ_{n,∞}, λ_{z,∞}) = 0
+$$
+
+Here we set $B_y(z=\infty) = 1/2 B_y(z=0)$ and $B_0 = 2 B_z$, and fix $λ_{z,∞} = -1$.
+By varying $λ_{n,∞}$, we find that the magnetic field profiles are exactly the same, while plasma velocity profiles vary.
+We normalize the plasma velocity by asympotic Alfvén velocity $v_{A,∞} = B_0 / \sqrt{n}$, and the profiles are plotted below @fig-uxnorm.
+It could be seen that for $λ_{n,∞} = 1$, we have zero bulk velocity change across the current sheet in the asymptotic limit.
+
+::: {.column-screen-inset}
+::: {#fig-profiles layout-nrow="1" }
+
+![Magnetic field, ion density, and out ion bulk velocity for $λ_{n,∞}=1$ and $λ_{z,∞}=-1$.](figures/profiles_ratio=0.5.svg){#fig-profiles_ratio}
+
+![Normalized Plasma Velocity $U_x$ Profiles for Different $λ_{n,∞}$.](figures/UxNorm.svg){#fig-uxnorm}
+
+![Normalized Plasma Velocity $U_y$ Profiles for Different $λ_{n,∞}$.](figures/UyNorm.svg){#fig-uynorm}
+
+:::
+:::
+
+![](figures/vRatios.svg)
+
 ## References
@@
 - @steinhauerMultifluidModelOnedimensional2008
+
+```mathematica
+conf = {ratio -> 1/2}
+TextString[conf]
+<!-- Convert a dict-like configuration to a string -->
+saveName[c_] := StringJoin[ToString /@ c]
+```

### 2024-07-11 index.qmd
add overleaf and abstract

@@
     href: zotero://select/library/collections/DZHNUFNV
+  - text: Overleaf
+    href: https://www.overleaf.com/project/668273f5d14067a22517615c
+abstract: "Solar wind plasma flow is filled by intense large-amplitude current sheets, so-called solar wind discontinuities, that are believed to be important regions for magnetic reconnection, solar wind ion heating, and energetic particle scattering.
+Although solar wind discontinuities share many properties of the classical MHD model of rotational discontinuity, their characteristics are more complicated and include many kinetic properties, going beyond such model.
+One of the most puzzling discontinuity properties is their Alvenicity: relation between Alfven and plasma velocity jumps across discontinuities.
+The high correlation of these jumps indicates the rotational nature of discontinuities, but the significant difference in jump amplitudes demonstrates that kinetic mechanisms are needed to describe the stress balance in discontinuities.
+Spacecraft observations demonstrate that plasma thermal anisotropy may not be sufficient to describe this difference and provide stress balance.
+In this study we show a theoretical model where this difference is explained by kinetics of interpenetrating ion beams crossing the discontinuity.
+The model successfully describes the observed discontinuity properties within the inner heliosphere."
 ---

### 2024-07-11 index.qmd
revise abstract (gpt)

--- 2024-06-07 index.qmd
@@
     href: zotero://select/library/collections/DZHNUFNV
+  - text: Overleaf
+    href: https://www.overleaf.com/project/668273f5d14067a22517615c
+abstract: "Solar wind plasma flow is characterized by intense, large-amplitude current sheets also known as solar wind discontinuities.
+These regions are believed to play a crucial role in magnetic reconnection, solar wind ion heating, and energetic particle scattering.
+Although solar wind discontinuities share several properties with the classical MHD model of rotational discontinuities, they exhibit more complex characteristics that include many kinetic properties, going beyond this model.
+One of the most puzzling properties of these discontinuities is their Alvenicity, which describes the relationship between Alfvén and plasma velocity jumps across discontinuities.
+While the high correlation of these jumps suggests the rotational nature of discontinuities, the significant differences in jump amplitudes indicate the necessity of kinetic mechanisms to describe the stress balance within discontinuities.
+Spacecraft observations show that plasma thermal anisotropy alone may not sufficiently explain this difference and achieve stress balance.
+In this study, we present a theoretical model where this difference is explained by the kinetics of interpenetrating ion beams crossing the discontinuity.
+Our model successfully describes the observed properties of discontinuities within the inner heliosphere."
 ---

### 2024-07-12 article.qmd
WIP: poster and paper

---
title: On a Alvenicity of solar wind discontinuities
authors:
  - Sergey Kamaletdinov
---

### 2024-07-12 index.qmd
WIP: poster and paper

@@
     href: https://www.overleaf.com/project/668273f5d14067a22517615c
-abstract: "Solar wind plasma flow is characterized by intense, large-amplitude current sheets also known as solar wind discontinuities.
+abstract: "Solar wind plasma flow is characterized by intense, large-amplitude current sheets, also known as solar wind discontinuities.
 These regions are believed to play a crucial role in magnetic reconnection, solar wind ion heating, and energetic particle scattering.
@@
 One of the most puzzling properties of these discontinuities is their Alvenicity, which describes the relationship between Alfvén and plasma velocity jumps across discontinuities.
-While the high correlation of these jumps suggests the rotational nature of discontinuities, the significant differences in jump amplitudes indicate the necessity of kinetic mechanisms to describe the stress balance within discontinuities.
+Although the high correlation of these jumps suggests the rotational nature of discontinuities, the significant differences in jump amplitudes indicate the necessity of kinetic mechanisms to describe the stress balance within discontinuities.
 Spacecraft observations show that plasma thermal anisotropy alone may not sufficiently explain this difference and achieve stress balance.
-In this study, we present a theoretical model where this difference is explained by the kinetics of interpenetrating ion beams crossing the discontinuity.
+In this study, we present a theoretical model in which this difference is explained by the kinetics of interpenetrating ion beams crossing the discontinuity.
 Our model successfully describes the observed properties of discontinuities within the inner heliosphere."
@@
 
+## Different normalization
+
+Normalized the quantities of the ion species
+
+$$
+n_α := λ_{αn} n_e
+\\
+u_{αx} := λ_{αu} B_z / \sqrt{n_{e,\infty}}
+\\
+u_{αy} := λ_{αu} B_z / \sqrt{n_{e,\infty}}
+\\
+u_{αz} := λ_{αu} B_z / \sqrt{n_{e,\infty}}
+$$
+
+$$
+θ' = \mp \frac{Λ_u}{B_0} + Γ_e
+$$
+
 ## References

### 2024-07-22 article.qmd
update: writing manuscript

@@
 authors:
+  - Zijin Zhang
+  - Xiaofei Shi
   - Sergey Kamaletdinov
+  - name: Anton V. Artemyev
+    email: aartemyev@igpp.ucla.edu
+    orcid: 0000-0001-8823-4474
+    affiliation: University of California, Los Angeles
+  - name: Vassilis Angelopoulos
+    orcid: 0000-0001-7024-1561
+    affiliation: University of California, Los Angeles
+code-links:
+  - text: Overleaf
+    href: https://www.overleaf.com/project/668273f5d14067a22517615c
+abstract: "Solar wind plasma flow is characterized by intense, large-amplitude current sheets, also known as solar wind discontinuities.
+These regions are believed to play a crucial role in magnetic reconnection, solar wind ion heating, and energetic particle scattering.
+Although solar wind discontinuities share several properties with the classical MHD model of rotational discontinuities, they exhibit more complex characteristics that include many kinetic properties, going beyond this model.
+One of the most puzzling properties of these discontinuities is their Alvenicity, which describes the relationship between Alfvén and plasma velocity jumps across discontinuities.
+Although the high correlation of these jumps suggests the rotational nature of discontinuities, the significant differences in jump amplitudes indicate the necessity of kinetic mechanisms to describe the stress balance within discontinuities.
+Spacecraft observations show that plasma thermal anisotropy alone may not sufficiently explain this difference and achieve stress balance.
+In this study, we present a theoretical model in which this difference is explained by the kinetics of interpenetrating ion beams crossing the discontinuity.
+Our model successfully describes the observed properties of discontinuities within the inner heliosphere."
 ---
+
+:::{.hide}
+\newcommand{\vect}[1]{\mathbf{#1}}
+:::
+
+# Introduction
+
+# Plasma-Field Equations for Force-free current sheet
+
+Following @steinhauerMultifluidModelOnedimensional2008, we consider a one-dimensional force-free current sheet where $B_x^2 + B_y^2 = B_0^2 = \text{const}$, with all variables dependent solely on $z$.
+The system comprises multiple steady-state ion species (fluid groups) and a background electron fluid.
+
+Electron motion is along the field lines $\vect{u_e} = Γ_e \vect{B} / (n_e B_z)$.
+And the conservation of fluid mass $d(n_α u_{αz})/dz = 0$ integrate to a constant parameters $Γ_α = n_α u_{αz}$.
+The momentum equations for each ion species are given by
+
+$$
+\begin{aligned}
+Γ_α \frac{d u_{α x}}{d z} &= n_α u_{α y} B_z - Γ_α B_y
+\\
+Γ_α \frac{d u_{α y}}{d z} &= Γ_α B_x-n_α u_{α x} B_z
+\\
+Γ_α \frac{d u_{α z}}{d z} &= -\frac{1}{2} \frac{d p_α}{d z}-n_α \frac{d \phi}{d z}+n_α u_{α x} B_y-n_α u_{α y} B_x
+\end{aligned}
+$${#eq-momentum}
+
+Ampere’s law connects the fields and the flow components
+
+$$
+d B_y / d z = - J_x = - \sum_α n_α u_{α x} + n_e u_{e x} = -n u_x+\Gamma_e B_x / B_z
+$${#eq-Jx}
+
+$$
+d B_x / d z = J_y = \sum_α n_α u_{α y} - n_e u_{e y} = n u_y-\Gamma_e B_y / B_z
+$${#eq-Jy}
+
+And the Gauss's law for magnetism gives another constant parameter $B_{z} = \text{const}$.
+
+This is a system of $3N+2$ equations for $4N+2$ unknown dependent variables: $B_x$, $B_y$, and $N$ each of $n$, $p$, $u_x$, and $u_y$.
+Since we are interested in force-free solutions with $B_x^2 + B_y^2 = B_0^2 = \text{const}$, this condition provides an additional equation for the system.
+
+A scalar pressure model, which provides another N equations connecting $p$ and $n$, was attempted as per @steinhauerMultifluidModelOnedimensional2008, but no solution was found (the system was overdetermined).
+Therefore, the system is considered with free parameters.
+For a system with two ion species, there are 9 equations for 10 unknowns.
+The system can be solved by setting one of the parameters to have a specific profile and solving the rest of the system.
+
+The force-free condition let us express the magnetic field in terms of the rotation angle $θ$ as $B_x = B_0 \cos θ$ and $B_y = B_0 \sin θ$. From Ampere's law (@eq-Jx) and (@eq-Jy), we have:
+
+$$
+n (u_x B_y  - u_y B_x) = 0
+$$
+
+And express the density and bulk velocity in terms of every ion species: $n_α u_{αx} B_y = n_α u_{αy} B_x + C_α$.
+This above equation can be rewritten as $\sum_α C_α =0$.
+The simplest case is to set $C_α=0$ for all species, which is the case we will consider in this study.
+
+$$
+n_α (u_{αx} B_y - n_α u_{αy} B_x) = 0
+$${#eq-alpha}
+
+Combining the above equation with the first two equations in (@eq-momentum), we have:
+
+$$
+u_{αx}^2 + u_{αy}^2 = const
+$$
+
+This would let us express the bulk velocity in terms of the rotation angle $θ_α$ as $u_{αx} = u_{α0} \cos θ_α$ and $u_{αy} = u_{α0} \sin θ_α$.
+And substituting the above expression into the our assumption (@eq-alpha), we have:
+
+$$
+\tan θ = \tan θ_α
+$$
+
+__#TODO#__ Here we consider the case where $θ_α = θ$ for all species, but the same procedure can be applied to the case where $θ_α = θ + \pi$ and it would give the same results.
+
+The first momentum equation (@eq-momentum) after substituting the above expression becomes:
+
+$$
+- u_α \sin θ_α θ_α' = n_α u_α \sin θ_α B_z / Γ_α - B_0 \sin θ
+$$
+
+$$
+\Rightarrow 
+θ_α' = - \frac{n_α B_z}{Γ_α} + \frac{B_0}{u_α}
+$${#eq-theta1}
+
+Note that $θ_α, n_α$ are dependent variables, and $u_α, B_0, B_z, Γ_α$ are constants determined by the system.
+So given the profile of $n_α$, we could solve the above equation to get the profile of $θ_α$, thus the profile of $u_α$.
+
+# Results
+
+For the same asymptotic magnetic field, it is interesting to see how the plasma profiles change with different system parameters.
+
+$$
+\text{cond}(n(z), λ_{n,∞}, λ_{z,∞}) = 0
+$$
+
+Here we set $B_y(z=\infty) = 1/2 B_y(z=0)$ and $B_0 = 2 B_z$, and fix $λ_{z,∞} = -1$.
+By varying $λ_{n,∞}$, we find that the magnetic field profiles are exactly the same, while plasma velocity profiles vary.
+We normalize the plasma velocity by asympotic Alfvén velocity $v_{A,∞} = B_0 / \sqrt{n}$, and the profiles are plotted below @fig-uxnorm.
+It could be seen that for $λ_{n,∞} = 1$, we have zero bulk velocity change across the current sheet in the asymptotic limit.
+
+![](figures/vRatios.svg)
+
+# References
+
+- @steinhauerMultifluidModelOnedimensional2008

### 2024-07-22 index.qmd

@@
     href: zotero://select/library/collections/DZHNUFNV
-  - text: Overleaf
-    href: https://www.overleaf.com/project/668273f5d14067a22517615c
-abstract: "Solar wind plasma flow is characterized by intense, large-amplitude current sheets, also known as solar wind discontinuities.
-These regions are believed to play a crucial role in magnetic reconnection, solar wind ion heating, and energetic particle scattering.
-Although solar wind discontinuities share several properties with the classical MHD model of rotational discontinuities, they exhibit more complex characteristics that include many kinetic properties, going beyond this model.
-One of the most puzzling properties of these discontinuities is their Alvenicity, which describes the relationship between Alfvén and plasma velocity jumps across discontinuities.
-Although the high correlation of these jumps suggests the rotational nature of discontinuities, the significant differences in jump amplitudes indicate the necessity of kinetic mechanisms to describe the stress balance within discontinuities.
-Spacecraft observations show that plasma thermal anisotropy alone may not sufficiently explain this difference and achieve stress balance.
-In this study, we present a theoretical model in which this difference is explained by the kinetics of interpenetrating ion beams crossing the discontinuity.
-Our model successfully describes the observed properties of discontinuities within the inner heliosphere."
 ---
@@
 
-The momentum equations (@eq-momentum) become:
+The first momentum equation (@eq-momentum) after substituting the above expression becomes:
 
 $$
-- u_1 \sin θ_1 θ_1' = n_1 u_1 \sin θ_1 B_z / Γ_1 - B_0 \sin θ
-$$$$
+- u_α \sin θ_α θ_α' = n_α u_α \sin θ_α B_z / Γ_α - B_0 \sin θ
+$$
+
+$$
 \Rightarrow 
-θ' = - \frac{n_1 B_z}{Γ_1} \pm \frac{B_0}{u_1}
+θ_α' = - \frac{n_α B_z}{Γ_α} \pm \frac{B_0}{u_α}
 $${#eq-theta1}
 
-- Note that $θ_1, n_1$ are dependent variables, and $u_1, B_0, B_z, Γ_1$ are constants determined by the system.
-- **Given $n_1$, we could solve $θ_1$ from the above equation.**
+- Note that $θ_α, n_α$ are dependent variables, and $u_α, B_0, B_z, Γ_α$ are constants determined by the system.
+- **So given the profile of $n_α$, we could solve the above equation to get the profile of $θ_α$, thus the profile of $u_α$.**
 - The derivative of $θ$ goes to zero at infinity, gives us a relation between $u_1$ and $B_0$:
@@
 
-## References
-
-- @steinhauerMultifluidModelOnedimensional2008
-
 ```mathematica

### 2024-08-09 article.qmd
update manuscripts and notebooks

@@
 
+$$
+d B_z / d z = J_z = \sum_α n_α u_{α z} - n_e u_{e z} = n u_z-\Gamma_e
+$${#eq-Jz}
+
 And the Gauss's law for magnetism gives another constant parameter $B_{z} = \text{const}$.
@@
 
-And express the density and bulk velocity in terms of every ion species: $n_α u_{αx} B_y = n_α u_{αy} B_x + C_α$.
+And we could express the density and bulk velocity in terms of every ion species: $n_α u_{αx} B_y = n_α u_{αy} B_x + C_α$.
 This above equation can be rewritten as $\sum_α C_α =0$.
@@
 
-This would let us express the bulk velocity in terms of the rotation angle $θ_α$ as $u_{αx} = u_{α0} \cos θ_α$ and $u_{αy} = u_{α0} \sin θ_α$.
+This would let us express the bulk velocity in terms of the rotation angle $θ_α$ as $u_{αx} = u_{α} \cos θ_α$ and $u_{αy} = u_{α} \sin θ_α$.
 And substituting the above expression into the our assumption (@eq-alpha), we have:
@@
 
-Note that $θ_α, n_α$ are dependent variables, and $u_α, B_0, B_z, Γ_α$ are constants determined by the system.
-So given the profile of $n_α$, we could solve the above equation to get the profile of $θ_α$, thus the profile of $u_α$.
+Note that $θ_α, n_α$ are dependent variables, and $u_α, B_0, B_z, Γ_α$ are constants pre-determined by the system.
+So given the profile of $n_α$ and the boundary condition, we could solve the above equation to get the profile of $θ_α$, thus the profile of $u_α$.
+
+The Ampere's law (@eq-Jy) after substitution becomes:
+
+$$
+- B_0 \sin θ θ' = \sum n_α u_{α} \sin θ_α - Γ_e B_0 \sin θ / B_z
+% Λ_y n_1 u_1 \sin θ_1 - Λ_z Γ_1 B_0 \sin θ / B_z
+$$
+
+$$
+\Rightarrow 
+θ' = - \frac{ \sum n_α u_{α} }{B_0} + \frac{Γ_e}{B_z}
+$${#eq-theta}
+
+The above equation relates the rotation angle $θ$ to the plasma bulk velocity $n u \equiv \sum n_α u_{α}$.
+By equating the above two equations of $θ'$, we could get a equation relating the plasma bulk velocity to one specific species:
+
+$$
+n u = B_0 \left(\frac{Γ_e}{B_z}+\frac{n_α B_z}{Γ_α}-\frac{B_0}{u_α}\right)
+$$
+
+In the asymptotic region, as all variables approach constant values, the derivatives must vanish.
+For @eq-theta1, this means $θ_α(∞)' = 0 = - \frac{n_α(∞) B_z}{Γ_α} + \frac{B_0}{u_α}$.
+This relates the velocity of each species to the asymptotic density.
+Rewriting the @eq-theta1 in terms of asymptotic values, we have:
+
+$$
+θ_α' = \frac{B_z}{Γ_α} (n_α - n_α(∞))
+$${#eq-theta1-asym}
+
+Conveniently, the center of the current sheet is chosen as the origin $z = 0$, which corresponds to the lower boundary of @eq-theta1.
+As a result, the boundary condition for the rotation angle is given by  $\theta_\alpha(0) = \pi/2$.
+
+Combine momentum equation (x) (@eq-momentum) and Ampere's law (y) (@eq-Jy) with condition that the constant of integration vanishes at the current sheet center yield
+
+$$
+B_x B_z = \sum n_α u_{α,x} u_{α,z} = \sum Γ_α u_{α,x}
+$$
+
+Evaluate the above equation in the asymptotic limit, we have
+
+$$
+{B_z}^2 = \sum Γ_{α}^2/n_{α}(∞)
+$${#eq-asym1}
+
+To simplify the analysis, it is useful to employ a dimensionless system by normalizing the variables with their asymptotic values: in this study, the magnetic field is normalized by $B_{\text{ref}} = B_z$ , and the density is normalized by $n_{\text{ref}} = n(∞)$.
+Other reference values derived from these two are displayed in the table below.
+
+| Variable | Reference Value |
+| --- | --- |
+| Frequency (ion plasma frequency) | $\omega_{pi} = \sqrt{\frac{n_{\text{ref}} e^2}{m_i \varepsilon_0}}$ |
+| Length (ion inertial length) | $L_{\text{ref}} = c / \omega_{pi}$ |
+| Velocity (Alfvén velocity) | $V_{\text{ref}} = B_{\text{ref}} / \sqrt{\mu_0 m_i n_{\text{ref}}}$ |
 
@@
 
+TODO: define and compare current sheet thickness
+
+TODO: compare with Harris current sheet solution with same thickness
+
+The multifluid model developed above is used here to study a specific example and the effects of model parameters: assume a density profile for one species taking the form of a Gaussian function
+
+$$
+n_1(z) \to n_1 (∞) (\frac{c_1}{\frac{z}{\delta_1 }^2+1}+1)
+$$
+
+Solving the @eq-theta1-asym analytically, we have the rotation angle profile as
+
+$$
+\theta (z)\to \frac{\pi  \Gamma _1-2 c_1 \delta _1 B_z n_1(\infty ) \tan ^{-1}\left(\frac{z}{\delta _1}\right)}{2 \Gamma _1}
+$$
+
+For the simplest situation where we have two group of ions of the same density but opposite bulk velocity, i.e., $n_1 = n_2, u_1 = -u_2$, the system could be determined given $c_1$, $δ_1$, and $B_0$.
+The system profile of magnetic field, plasma density, and plasma velocity for one specific case ($δ_1=1, c_1=1/\sqrt{2}, B_0 = 2$) is plotted below @fig-profiles.
+The specific profiles are chosen to have zero electron current across the current sheet and zero $B_y$ in the asymptotic limit, corresponding to a 180° rotation of the magnetic field across the current sheet.
+However, in general, we would expect the electron current to be non-zero.
+
+::: {.column-page}
+::: {#fig-profiles layout-nrow="1" }
+
+![](figures/profiles_sym.svg)
+
+![](figures/J_profiles_sym.svg)
+
+(a). Magnetic field, ion density, and ion bulk velocity for $n_1 = n_2, u_1 = -u_2$ and $δ_1=1, c_1=1/\sqrt{2}, B_0 = 2$. (b). Current density profiles for the same case.
+:::
+:::
+
 For the same asymptotic magnetic field, it is interesting to see how the plasma profiles change with different system parameters.
 
-$$
-\text{cond}(n(z), λ_{n,∞}, λ_{z,∞}) = 0
-$$
+Here we set $B_y(z=\infty) = 1/2 B_0, B_0 = 2 B_z$, and let two ion species have the opposite bulk velocity but this time with different densities.
+By varying $n_1 (∞)$, we find that the magnetic field profiles are exactly the same, while plasma velocity profiles vary.
 
-Here we set $B_y(z=\infty) = 1/2 B_y(z=0)$ and $B_0 = 2 B_z$, and fix $λ_{z,∞} = -1$.
-By varying $λ_{n,∞}$, we find that the magnetic field profiles are exactly the same, while plasma velocity profiles vary.
 We normalize the plasma velocity by asympotic Alfvén velocity $v_{A,∞} = B_0 / \sqrt{n}$, and the profiles are plotted below @fig-uxnorm.
 It could be seen that for $λ_{n,∞} = 1$, we have zero bulk velocity change across the current sheet in the asymptotic limit.
+
+::: {.column-screen-inset}
+::: {#fig-profiles layout-nrow="1" }
+
+![Magnetic field, ion density, and ion bulk velocity for $λ_{n,∞}=1$ and $λ_{z,∞}=-1$.](figures/profiles_ratio=0.5.svg){#fig-profiles_ratio}
+
+![Normalized Plasma Velocity $U_x$ Profiles for Different $λ_{n,∞}$.](figures/UxNorm.svg){#fig-uxnorm}
+
+![Normalized Plasma Velocity $U_y$ Profiles for Different $λ_{n,∞}$.](figures/UyNorm.svg){#fig-uynorm}
+
+:::
+:::
 

### 2024-08-11 article.qmd
update: writing

@@
 
-TODO: define and compare current sheet thickness
-
-TODO: compare with Harris current sheet solution with same thickness
-
 The multifluid model developed above is used here to study a specific example and the effects of model parameters: assume a density profile for one species taking the form of a Gaussian function
@@
 
+## Alfvenicity
+
 For the same asymptotic magnetic field, it is interesting to see how the plasma profiles change with different system parameters.
@@
 Here we set $B_y(z=\infty) = 1/2 B_0, B_0 = 2 B_z$, and let two ion species have the opposite bulk velocity but this time with different densities.
-By varying $n_1 (∞)$, we find that the magnetic field profiles are exactly the same, while plasma velocity profiles vary.
+By varying $n_1(∞)$, we find that the magnetic field profiles are exactly the same, while plasma density and velocity  profiles vary.
 
-We normalize the plasma velocity by asympotic Alfvén velocity $v_{A,∞} = B_0 / \sqrt{n}$, and the profiles are plotted below @fig-uxnorm.
-It could be seen that for $λ_{n,∞} = 1$, we have zero bulk velocity change across the current sheet in the asymptotic limit.
+We normalize the plasma velocity by asympotic Alfvén velocity $v_{A}(∞) = B_0 / \sqrt{\mu_0 m_i n}$, and the profiles are plotted below @fig-profilesC.
+It could be seen that for $n_1(∞) = 0.5$, we have zero bulk velocity change across the current sheet in the asymptotic limit.
+And the change of bulk velocity across the current sheet increases with $n_1(∞)$ decreasing to zero @fig-uxnorm.
+The normalized plasma velocity $U_y$ in the aympotic limit decreases with $n_1(∞)$ decreasing to zero @fig-uynorm.
 
-::: {.column-screen-inset}
-::: {#fig-profiles layout-nrow="1" }
+::: {.column-page}
+::: {#fig-profilesC layout-nrow="1" }
 
-![Magnetic field, ion density, and ion bulk velocity for $λ_{n,∞}=1$ and $λ_{z,∞}=-1$.](figures/profiles_ratio=0.5.svg){#fig-profiles_ratio}
+![Magnetic field, ion density, and ion bulk velocity for two ion species with same bulk velocity but different densities.](figures/profiles_ratio=0.5.svg){#fig-profiles_ratio}
 
-![Normalized Plasma Velocity $U_x$ Profiles for Different $λ_{n,∞}$.](figures/UxNorm.svg){#fig-uxnorm}
+![Plasma velocity $U_x$ profiles normalized by aysmpotic Alfvén velocity for different $n_1(∞)$.](figures/UxNormB0.svg){#fig-uxnorm}
 
-![Normalized Plasma Velocity $U_y$ Profiles for Different $λ_{n,∞}$.](figures/UyNorm.svg){#fig-uynorm}
+![Plasma velocity $U_y$ profiles normalized by aysmpotic Alfvén velocity for different $n_1(∞)$.](figures/UyNormB0.svg){#fig-uynorm}
 
@@
 
-![](figures/vRatios.svg)
+Normalized by Alfvén velocity $v_{A,x}$ and $v_{A,y}$, the plasma velocity profiles are plotted below @fig-uxNormBx and @fig-uyNormBy.
+The normalized plasma velocity profiles are exactly the same for $U_x/v_{A,x}$ and $U_y/v_{A,y}$ for same $n_1(∞)$.
+When $n_1(∞)$ decreases to zero, the normalized plasma velocity would approach one in the asymptotic limit, indicating the Alvenicity of the current sheet.
+
+::: {.column-page}
+::: {#fig-uProfilesNorm layout-nrow="1" }
+![Plasma velocity $U_x$ profiles normalized by $v_{A,x}$ for different $n_1(∞)$.](figures/UxNormBx.svg){#fig-uxNormBx}
+
+![Plasma velocity $U_y$ profiles normalized by $v_{A,y}$ for different $n_1(∞)$.](figures/UyNormBy.svg){#fig-uyNormBy}
+:::
+:::
+
+The asymptotic velocity for the each species, bulk velocity and normalized one by Alfvén velocity are plotted below @fig-vRatios and @fig-vRatiosEx2 for the case of $\lambda_{z,2} = - 1$ (two groups of opposite velocity in z direction) and $\lambda_{z,2} = - 1/2$ respectively.
+For the case of $\lambda_{z,2} = - 1$ as shown in @fig-vRatiosEx1, the normalized velocity reaches zero when $n_1(∞) = 0.5$.
+The bulk velocity of each species is the same in the asymptotic limit and does not depend on $n_1(∞)$.
+However, this does not generally hold when $\lambda_{z,2} \neq - 1$ as shown in @fig-vRatiosEx2 for the case of $\lambda_{z,2} = - 1/2$..
+The normalized velocity reaches zero when $n_1(∞) = 1/3$, and the bulk velocity of each species differs in the asymptotic limit and depends on $n_1(∞)$.
+
+::: {.column-page}
+::: {#fig-vRatios layout-nrow="1" }
+![](figures/vRatios.svg){#fig-vRatiosEx1}
+
+![](figures/vRatiosEx2.svg){#fig-vRatiosEx2}
+
+Asymptotic velocity for each species, bulk velocity and normalized one by Alfvén velocity for the case of $\lambda_{z,2} = - 1$ (a) and $\lambda_{z,2} = - 1/2$ (b).
+:::
+:::
+
+## Current Sheet Thickness
+
+TODO: define and compare current sheet thickness
+
+TODO: compare with Harris current sheet solution with same thickness
+
+## Density Peaking
+
+# Summary
 

### 2024-08-12 article.qmd
updates

@@
 # Introduction
+
+1. Observation features of solar wind discontinuities
+2. The necessity of kinetic models to describe solar wind discontinuities
+3. Current theoretical models of solar wind discontinuities and their limitations
+  - Harris current sheet @harrisPlasmaSheathSeparating1962
+
+Force-free @harrisonRemarksOnedimensionalForcefree2009 (A necessary condition on the pseudopotential plasma pressure to allow for force-free 1D VM equilibria is formulated.)
+
+<!-- Work on self-consistent kinetic models of the current sheet was pioneered by Sestero [1966] and by Schindler [1972, 1975, 1979] who presumed a much larger scale in the plane of the current sheet than in the normal (z) direction.
+Kan [1973] obtained exact solutions for the two dimensional (2-D) tail using a single Maxwellian distribution having uniform drift velocity and also a two component Maxwellian distribution having a sheared drift.
+His technique has been used by others to obtain 2-D solutions with an X-point (reviewed by Yoon and Lui [2005]).
+Eastwood [1972], Hill [1975], Cowley [1978a, 1978b], Cowley and Pellat [1979], Francfort and Pellat [1976], and others have discussed both the fluid and kinetic aspects of the current sheet, showing the importance of the balance between magnetic field pressure and the particle pressure component along the magnetic field, pk.
+-->
+
+# Multifluid Model
+
+We used a steady-state multi-fluid model for assumed 1-D force-free current sheets.
+This approach, inspired by @steinhauerMultifluidModelOnedimensional2008, has the advantage of being analytically tractable while providing the necessary complexity to include the kinetic effects of multiple ion species.
+Kinetic approach, which is in principle more accurate, is hard to solve analytically and difficult to interpret the measurements.
+The multifluid model is a good compromise between the simplicity of the MHD model and the complexity of the kinetic model.
+One disadvantage of the multifluid model, however, is that there are more unknowns than equations.
+This means that the model is underdetermined and requires additional assumptions to be solved.
+This closure problem is usually solved by choosing equations of state.
+In this study, we first attempted to solve the system with a scalar pressure model, but then the system was overdetermined and no solution was found.
+We avoid this problem by assuming a specific profile for one of the variables, which is then used to solve the rest of the system.
+
+<!-- An example is the simultaneous appearance of separate beams of ions in different directions with different fluxes.
+Parks et al.
+[2001] and Parks [2004] have pointed out that microphysical processes are key to understanding these new observations that, evidently, are unexplainable by conventional magnetohydrodynamic models.
+-->
 

### 2024-08-14 article.qmd
update figures

@@
   - Harris current sheet @harrisPlasmaSheathSeparating1962
+
+Observations of magnetic discontinuities exhibit Alfvenic properties, with the plasma velocity variation correlating well with magnetic field components variations.
+However, the amplitude of the plasma velocity jump is often significantly smaller than the Alfvén velocity jump, indicating a more complicated picture than the classical MHD model of rotational discontinuities.
+@fig-examples show observations of  magnetic discontinuities in different regions of the heliosphere, PSP within 0.2 AU, Wind upstream of the Earth's bow shock, and ARTEMIS in the Earth's magnetotail.
+These differences could not be simply explained by the aniostropy of the plasma thermal pressure alone and suggest that kinetic mechanisms are necessary to describe the stress balance within discontinuities.
+
+::: {.column-page}
+::: {#fig-examples layout-nrow="1" }
+
+![]()
+
+Three examples of solar wind discontinuities observed by Parker Solar Probe (PSP), Wind and ARTEMIS spacecraft.
+
+:::
+:::
 
@@
 
+::: {#fig-vRatioStats}
+![](https://raw.githubusercontent.com/Beforerr/psp_conjunction/5e09973ebedbdaef5cc89cb3a12be3fb77f44570/figures/enc7/vl_ratio.png)
+
+Statistics of the asymptotic velocity ratio from PSP, Wind, and ARTEMIS spacecraft observations during PSP encounter 7 period from 2021-01-14 to 2021-01-21.
+:::
+
 # References
@@
 - @steinhauerMultifluidModelOnedimensional2008
+- @artemyevIonNongyrotropySolar2020

### 2024-08-16 article.qmd
update: justfile

@@
 
+# References
+
+- @steinhauerMultifluidModelOnedimensional2008
+- @artemyevIonNongyrotropySolar2020
+- @neukirchKineticModelsTangential2020
+- @vaskoThinCurrentSheets2014
+
+<!-- "This difference has previously been interpreted as: (1) a possible contribution of anisotropy that decreases ΔvA, or (2) a discontinuity non-stationarity due to residual magnetic energy.
+We propose an alternate interpretation: an ion nonadiabatic interaction with intense (thin) discontinuities that shapes the nongyrotropic ion distribution to include a nondiagonal term of the pressure tensor, with a cross-discontinuity gradient decreasing ΔvA." - @artemyevIonNongyrotropySolar2020 -->
+
 # Introduction
@@
 @fig-examples show observations of  magnetic discontinuities in different regions of the heliosphere, PSP within 0.2 AU, Wind upstream of the Earth's bow shock, and ARTEMIS in the Earth's magnetotail.
-These differences could not be simply explained by the aniostropy of the plasma thermal pressure alone and suggest that kinetic mechanisms are necessary to describe the stress balance within discontinuities.
+Previously, this discrepancy has been interpreted as a possible contribution of anisotropy that decreases the Alfvén velocity jump, or a discontinuity non-stationarity due to residual magnetic energy, or an ion nonadiabatic interaction with intense (thin) discontinuities that shapes the nongyrotropic ion distribution to include a nondiagonal term of the pressure tensor, with a cross-discontinuity gradient decreasing the Alfvén velocity jump @artemyevIonNongyrotropySolar2020.
+
+These differences could not be simply explained by the aniostropy of the plasma thermal pressure alone (assuming a Maxwellian distribution that fits both the parallel and perpendicular temperatures) and suggest that kinetic mechanisms are necessary to describe the stress balance within discontinuities.
 
@@
 ::: {#fig-examples layout-nrow="1" }
-
-![]()
 
@@
 :::
-
-# References
-
-- @steinhauerMultifluidModelOnedimensional2008
-- @artemyevIonNongyrotropySolar2020

### 2024-08-20 article.qmd
new: add examples figures

@@
 authors:
-  - Zijin Zhang
-  - Xiaofei Shi
-  - Sergey Kamaletdinov
+  - name: Zijin Zhang
+    email: zijin@ucla.edu
+    orcid: 0000-0002-9968-067X
+    affiliation: University of California, Los Angeles
+  - name: Sergei Kamaletdinov
+    email: sergei2033@ucla.edu
+    orcid: 0000-0002-2261-0331
+    affiliation: University of California, Los Angeles
+  - name: Xiaofei Shi
+    email: sxf1698@ucla.edu
+    orcid: 0000-0003-3367-5074
+    affiliation: University of California, Los Angeles
   - name: Anton V. Artemyev
@@
     orcid: 0000-0001-8823-4474
+    affiliation: University of California, Los Angeles
+  - name: Marco Velli
+    email: mvelli@ucla.edu
+    orcid: 0000-0002-2381-3106
     affiliation: University of California, Los Angeles

### 2024-08-20 _manuscript/_tex/article.tex
Built site for gh-pages

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{plainnat}
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={On a Alvenicity of solar wind discontinuities},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{On a Alvenicity of solar wind discontinuities}
\author{Zijin Zhang \and Sergei Kamaletdinov \and Xiaofei Shi \and Anton V. Artemyev \and Marco Velli \and Vassilis Angelopoulos}
\date{}

\begin{document}
\maketitle
\begin{abstract}
Solar wind plasma flow is characterized by intense, large-amplitude current sheets, also known as solar wind discontinuities. These regions are believed to play a crucial role in magnetic reconnection, solar wind ion heating, and energetic particle scattering. Although solar wind discontinuities share several properties with the classical MHD model of rotational discontinuities, they exhibit more complex characteristics that include many kinetic properties, going beyond this model. One of the most puzzling properties of these discontinuities is their Alvenicity, which describes the relationship between Alfvén and plasma velocity jumps across discontinuities. Although the high correlation of these jumps suggests the rotational nature of discontinuities, the significant differences in jump amplitudes indicate the necessity of kinetic mechanisms to describe the stress balance within discontinuities. Spacecraft observations show that plasma thermal anisotropy alone may not sufficiently explain this difference and achieve stress balance. In this study, we present a theoretical model in which this difference is explained by the kinetics of interpenetrating ion beams crossing the discontinuity. Our model successfully describes the observed properties of discontinuities within the inner heliosphere.
\end{abstract}

\newcommand{\vect}[1]{\mathbf{#1}}

\section{References}\label{references}

\begin{itemize}
\tightlist
\item
  \citet{steinhauerMultifluidModelOnedimensional2008}
\item
  \citet{artemyevIonNongyrotropySolar2020}
\item
  \citet{neukirchKineticModelsTangential2020}
\item
  \citet{vaskoThinCurrentSheets2014}
\end{itemize}

\section{Introduction}\label{introduction}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Observation features of solar wind discontinuities
\item
  The necessity of kinetic models to describe solar wind discontinuities
\item
  Current theoretical models of solar wind discontinuities and their limitations
\end{enumerate}

\begin{itemize}
\tightlist
\item
  Harris current sheet \citet{harrisPlasmaSheathSeparating1962}
\end{itemize}

Observations of magnetic discontinuities exhibit Alfvenic properties, with the plasma velocity variation correlating well with magnetic field components variations. However, the amplitude of the plasma velocity jump is often significantly smaller than the Alfvén velocity jump, indicating a more complicated picture than the classical MHD model of rotational discontinuities. Figure~\ref{fig-examples} show observations of magnetic discontinuities in different regions of the heliosphere, PSP within 0.2 AU, Wind upstream of the Earth's bow shock, and ARTEMIS in the Earth's magnetotail. Previously, this discrepancy has been interpreted as a possible contribution of anisotropy that decreases the Alfvén velocity jump, or a discontinuity non-stationarity due to residual magnetic energy, or an ion nonadiabatic interaction with intense (thin) discontinuities that shapes the nongyrotropic ion distribution to include a nondiagonal term of the pressure tensor, with a cross-discontinuity gradient decreasing the Alfvén velocity jump \citet{artemyevIonNongyrotropySolar2020}.

These differences could not be simply explained by the aniostropy of the plasma thermal pressure alone (assuming a Maxwellian distribution that fits both the parallel and perpendicular temperatures) and suggest that kinetic mechanisms are necessary to describe the stress balance within discontinuities.

\begin{figure}

\centering{

\includegraphics{article_files/mediabag/figures/fig_examples.pdf}

}

\caption{\label{fig-examples}Three examples of solar wind discontinuities observed by Parker Solar Probe (PSP), ARTEMIS and Wind spacecraft with sub-Alfvenic flow velocity.}

\end{figure}%

Force-free \citet{harrisonRemarksOnedimensionalForcefree2009} (A necessary condition on the pseudopotential plasma pressure to allow for force-free 1D VM equilibria is formulated.)

\section{Multifluid Model}\label{multifluid-model}

We used a steady-state multi-fluid model for assumed 1-D force-free current sheets. This approach, inspired by \citet{steinhauerMultifluidModelOnedimensional2008}, has the advantage of being analytically tractable while providing the necessary complexity to include the kinetic effects of multiple ion species. Kinetic approach, which is in principle more accurate, is hard to solve analytically and difficult to interpret the measurements. The multifluid model is a good compromise between the simplicity of the MHD model and the complexity of the kinetic model. One disadvantage of the multifluid model, however, is that there are more unknowns than equations. This means that the model is underdetermined and requires additional assumptions to be solved. This closure problem is usually solved by choosing equations of state. In this study, we first attempted to solve the system with a scalar pressure model, but then the system was overdetermined and no solution was found. We avoid this problem by assuming a specific profile for one of the variables, which is then used to solve the rest of the system.

\section{Plasma-Field Equations for Force-free current sheet}\label{plasma-field-equations-for-force-free-current-sheet}

Following \citet{steinhauerMultifluidModelOnedimensional2008}, we consider a one-dimensional force-free current sheet where \(B_x^2 + B_y^2 = B_0^2 = \text{const}\), with all variables dependent solely on \(z\). The system comprises multiple steady-state ion species (fluid groups) and a background electron fluid.

Electron motion is along the field lines \(\mathbf{u_e} = Γ_e \mathbf{B} / (n_e B_z)\). And the conservation of fluid mass \(d(n_α u_{αz})/dz = 0\) integrate to a constant parameters \(Γ_α = n_α u_{αz}\). The momentum equations for each ion species are given by

\begin{equation}\phantomsection\label{eq-momentum}{
\begin{aligned}
Γ_α \frac{d u_{α x}}{d z} &= n_α u_{α y} B_z - Γ_α B_y
\\
Γ_α \frac{d u_{α y}}{d z} &= Γ_α B_x-n_α u_{α x} B_z
\\
Γ_α \frac{d u_{α z}}{d z} &= -\frac{1}{2} \frac{d p_α}{d z}-n_α \frac{d \phi}{d z}+n_α u_{α x} B_y-n_α u_{α y} B_x
\end{aligned}
}\end{equation}

Ampere's law connects the fields and the flow components

\begin{equation}\phantomsection\label{eq-Jx}{
d B_y / d z = - J_x = - \sum_α n_α u_{α x} + n_e u_{e x} = -n u_x+\Gamma_e B_x / B_z
}\end{equation}

\begin{equation}\phantomsection\label{eq-Jy}{
d B_x / d z = J_y = \sum_α n_α u_{α y} - n_e u_{e y} = n u_y-\Gamma_e B_y / B_z
}\end{equation}

\begin{equation}\phantomsection\label{eq-Jz}{
d B_z / d z = J_z = \sum_α n_α u_{α z} - n_e u_{e z} = n u_z-\Gamma_e
}\end{equation}

And the Gauss's law for magnetism gives another constant parameter \(B_{z} = \text{const}\).

This is a system of \(3N+2\) equations for \(4N+2\) unknown dependent variables: \(B_x\), \(B_y\), and \(N\) each of \(n\), \(p\), \(u_x\), and \(u_y\). Since we are interested in force-free solutions with \(B_x^2 + B_y^2 = B_0^2 = \text{const}\), this condition provides an additional equation for the system.

A scalar pressure model, which provides another N equations connecting \(p\) and \(n\), was attempted as per \citet{steinhauerMultifluidModelOnedimensional2008}, but no solution was found (the system was overdetermined). Therefore, the system is considered with free parameters. For a system with two ion species, there are 9 equations for 10 unknowns. The system can be solved by setting one of the parameters to have a specific profile and solving the rest of the system.

The force-free condition let us express the magnetic field in terms of the rotation angle \(θ\) as \(B_x = B_0 \cos θ\) and \(B_y = B_0 \sin θ\). From Ampere's law (Equation~\ref{eq-Jx}) and (Equation~\ref{eq-Jy}), we have:

\[
n (u_x B_y  - u_y B_x) = 0
\]

And we could express the density and bulk velocity in terms of every ion species: \(n_α u_{αx} B_y = n_α u_{αy} B_x + C_α\). This above equation can be rewritten as \(\sum_α C_α =0\). The simplest case is to set \(C_α=0\) for all species, which is the case we will consider in this study.

\begin{equation}\phantomsection\label{eq-alpha}{
n_α (u_{αx} B_y - n_α u_{αy} B_x) = 0
}\end{equation}

Combining the above equation with the first two equations in (Equation~\ref{eq-momentum}), we have:

\[
u_{αx}^2 + u_{αy}^2 = const
\]

This would let us express the bulk velocity in terms of the rotation angle \(θ_α\) as \(u_{αx} = u_{α} \cos θ_α\) and \(u_{αy} = u_{α} \sin θ_α\). And substituting the above expression into the our assumption (Equation~\ref{eq-alpha}), we have:

\[
\tan θ = \tan θ_α
\]

\textbf{\#TODO\#} Here we consider the case where \(θ_α = θ\) for all species, but the same procedure can be applied to the case where \(θ_α = θ + \pi\) and it would give the same results.

The first momentum equation (Equation~\ref{eq-momentum}) after substituting the above expression becomes:

\[
- u_α \sin θ_α θ_α' = n_α u_α \sin θ_α B_z / Γ_α - B_0 \sin θ
\]

\begin{equation}\phantomsection\label{eq-theta1}{
\Rightarrow 
θ_α' = - \frac{n_α B_z}{Γ_α} + \frac{B_0}{u_α}
}\end{equation}

Note that \(θ_α, n_α\) are dependent variables, and \(u_α, B_0, B_z, Γ_α\) are constants pre-determined by the system. So given the profile of \(n_α\) and the boundary condition, we could solve the above equation to get the profile of \(θ_α\), thus the profile of \(u_α\).

The Ampere's law (Equation~\ref{eq-Jy}) after substitution becomes:

\[
- B_0 \sin θ θ' = \sum n_α u_{α} \sin θ_α - Γ_e B_0 \sin θ / B_z
% Λ_y n_1 u_1 \sin θ_1 - Λ_z Γ_1 B_0 \sin θ / B_z
\]

\begin{equation}\phantomsection\label{eq-theta}{
\Rightarrow 
θ' = - \frac{ \sum n_α u_{α} }{B_0} + \frac{Γ_e}{B_z}
}\end{equation}

The above equation relates the rotation angle \(θ\) to the plasma bulk velocity \(n u \equiv \sum n_α u_{α}\). By equating the above two equations of \(θ'\), we could get a equation relating the plasma bulk velocity to one specific species:

\[
n u = B_0 \left(\frac{Γ_e}{B_z}+\frac{n_α B_z}{Γ_α}-\frac{B_0}{u_α}\right)
\]

In the asymptotic region, as all variables approach constant values, the derivatives must vanish. For Equation~\ref{eq-theta1}, this means \(θ_α(∞)' = 0 = - \frac{n_α(∞) B_z}{Γ_α} + \frac{B_0}{u_α}\). This relates the velocity of each species to the asymptotic density. Rewriting the Equation~\ref{eq-theta1} in terms of asymptotic values, we have:

\begin{equation}\phantomsection\label{eq-theta1-asym}{
θ_α' = \frac{B_z}{Γ_α} (n_α - n_α(∞))
}\end{equation}

Conveniently, the center of the current sheet is chosen as the origin \(z = 0\), which corresponds to the lower boundary of Equation~\ref{eq-theta1}. As a result, the boundary condition for the rotation angle is given by \(\theta_\alpha(0) = \pi/2\).

Combine momentum equation (x) (Equation~\ref{eq-momentum}) and Ampere's law (y) (Equation~\ref{eq-Jy}) with condition that the constant of integration vanishes at the current sheet center yield

\[
B_x B_z = \sum n_α u_{α,x} u_{α,z} = \sum Γ_α u_{α,x}
\]

Evaluate the above equation in the asymptotic limit, we have

\begin{equation}\phantomsection\label{eq-asym1}{
{B_z}^2 = \sum Γ_{α}^2/n_{α}(∞)
}\end{equation}

To simplify the analysis, it is useful to employ a dimensionless system by normalizing the variables with their asymptotic values: in this study, the magnetic field is normalized by \(B_{\text{ref}} = B_z\) , and the density is normalized by \(n_{\text{ref}} = n(∞)\). Other reference values derived from these two are displayed in the table below.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.5000}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.5000}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Variable
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Reference Value
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Frequency (ion plasma frequency) & \(\omega_{pi} = \sqrt{\frac{n_{\text{ref}} e^2}{m_i \varepsilon_0}}\) \\
Length (ion inertial length) & \(L_{\text{ref}} = c / \omega_{pi}\) \\
Velocity (Alfvén velocity) & \(V_{\text{ref}} = B_{\text{ref}} / \sqrt{\mu_0 m_i n_{\text{ref}}}\) \\
\end{longtable}

\section{Results}\label{results}

The multifluid model developed above is used here to study a specific example and the effects of model parameters: assume a density profile for one species taking the form of a Gaussian function

\[
n_1(z) \to n_1 (∞) (\frac{c_1}{\frac{z}{\delta_1 }^2+1}+1)
\]

Solving the Equation~\ref{eq-theta1-asym} analytically, we have the rotation angle profile as

\[
\theta (z)\to \frac{\pi  \Gamma _1-2 c_1 \delta _1 B_z n_1(\infty ) \tan ^{-1}\left(\frac{z}{\delta _1}\right)}{2 \Gamma _1}
\]

For the simplest situation where we have two group of ions of the same density but opposite bulk velocity, i.e., \(n_1 = n_2, u_1 = -u_2\), the system could be determined given \(c_1\), \(δ_1\), and \(B_0\). The system profile of magnetic field, plasma density, and plasma velocity for one specific case (\(δ_1=1, c_1=1/\sqrt{2}, B_0 = 2\)) is plotted below Figure~\ref{fig-profiles}. The specific profiles are chosen to have zero electron current across the current sheet and zero \(B_y\) in the asymptotic limit, corresponding to a 180° rotation of the magnetic field across the current sheet. However, in general, we would expect the electron current to be non-zero.

\begin{figure}

\begin{minipage}{0.50\linewidth}
\includegraphics{article_files/mediabag/figures/profiles_sym.pdf}\end{minipage}%
%
\begin{minipage}{0.50\linewidth}
\includegraphics{article_files/mediabag/figures/J_profiles_sym.pdf}\end{minipage}%

\caption{\label{fig-profiles}(a). Magnetic field, ion density, and ion bulk velocity for \(n_1 = n_2, u_1 = -u_2\) and \(δ_1=1, c_1=1/\sqrt{2}, B_0 = 2\). (b). Current density profiles for the same case.}

\end{figure}%

\subsection{Alfvenicity}\label{alfvenicity}

For the same asymptotic magnetic field, it is interesting to see how the plasma profiles change with different system parameters.

Here we set \(B_y(z=\infty) = 1/2 B_0, B_0 = 2 B_z\), and let two ion species have the opposite bulk velocity but this time with different densities. By varying \(n_1(∞)\), we find that the magnetic field profiles are exactly the same, while plasma density and velocity profiles vary.

We normalize the plasma velocity by asympotic Alfvén velocity \(v_{A}(∞) = B_0 / \sqrt{\mu_0 m_i n}\), and the profiles are plotted below Figure~\ref{fig-profilesC}. It could be seen that for \(n_1(∞) = 0.5\), we have zero bulk velocity change across the current sheet in the asymptotic limit. And the change of bulk velocity across the current sheet increases with \(n_1(∞)\) decreasing to zero Figure~\ref{fig-uxnorm}. The normalized plasma velocity \(U_y\) in the aympotic limit decreases with \(n_1(∞)\) decreasing to zero Figure~\ref{fig-uynorm}.

\begin{figure}

\begin{minipage}{0.25\linewidth}

\centering{

\includegraphics{article_files/mediabag/figures/profiles_ratio=0.5.pdf}

}

\subcaption{\label{fig-profiles_ratio}Magnetic field, ion density, and ion bulk velocity for two ion species with same bulk velocity but different densities.}

\end{minipage}%
%
\begin{minipage}{0.25\linewidth}

\centering{

\includegraphics{article_files/mediabag/figures/UxNormB0.pdf}

}

\subcaption{\label{fig-uxnorm}Plasma velocity \(U_x\) profiles normalized by aysmpotic Alfvén velocity for different \(n_1(∞)\).}

\end{minipage}%
%
\begin{minipage}{0.25\linewidth}

\centering{

\includegraphics{article_files/mediabag/figures/UyNormB0.pdf}

}

\subcaption{\label{fig-uynorm}Plasma velocity \(U_y\) profiles normalized by aysmpotic Alfvén velocity for different \(n_1(∞)\).}

\end{minipage}%
%
\begin{minipage}{0.25\linewidth}

\end{minipage}%

\end{figure}%

Normalized by Alfvén velocity \(v_{A,x}\) and \(v_{A,y}\), the plasma velocity profiles are plotted below Figure~\ref{fig-uxNormBx} and Figure~\ref{fig-uyNormBy}. The normalized plasma velocity profiles are exactly the same for \(U_x/v_{A,x}\) and \(U_y/v_{A,y}\) for same \(n_1(∞)\). When \(n_1(∞)\) decreases to zero, the normalized plasma velocity would approach one in the asymptotic limit, indicating the Alvenicity of the current sheet.

\begin{figure}

\begin{minipage}{0.50\linewidth}

\centering{

\includegraphics{article_files/mediabag/figures/UxNormBx.pdf}

}

\subcaption{\label{fig-uxNormBx}Plasma velocity \(U_x\) profiles normalized by \(v_{A,x}\) for different \(n_1(∞)\).}

\end{minipage}%
%
\begin{minipage}{0.50\linewidth}

\centering{

\includegraphics{article_files/mediabag/figures/UyNormBy.pdf}

}

\subcaption{\label{fig-uyNormBy}Plasma velocity \(U_y\) profiles normalized by \(v_{A,y}\) for different \(n_1(∞)\).}

\end{minipage}%

\end{figure}%

The asymptotic velocity for the each species, bulk velocity and normalized one by Alfvén velocity are plotted below Figure~\ref{fig-vRatios} and Figure~\ref{fig-vRatiosEx2} for the case of \(\lambda_{z,2} = - 1\) (two groups of opposite velocity in z direction) and \(\lambda_{z,2} = - 1/2\) respectively. For the case of \(\lambda_{z,2} = - 1\) as shown in Figure~\ref{fig-vRatiosEx1}, the normalized velocity reaches zero when \(n_1(∞) = 0.5\). The bulk velocity of each species is the same in the asymptotic limit and does not depend on \(n_1(∞)\). However, this does not generally hold when \(\lambda_{z,2} \neq - 1\) as shown in Figure~\ref{fig-vRatiosEx2} for the case of \(\lambda_{z,2} = - 1/2\).. The normalized velocity reaches zero when \(n_1(∞) = 1/3\), and the bulk velocity of each species differs in the asymptotic limit and depends on \(n_1(∞)\).

\begin{figure}

\begin{minipage}{0.50\linewidth}

\centering{

\captionsetup{labelsep=none}\includegraphics{article_files/mediabag/figures/vRatios.pdf}

}

\subcaption{\label{fig-vRatiosEx1}}

\end{minipage}%
%
\begin{minipage}{0.50\linewidth}

\centering{

\captionsetup{labelsep=none}\includegraphics{article_files/mediabag/figures/vRatiosEx2.pdf}

}

\subcaption{\label{fig-vRatiosEx2}}

\end{minipage}%

\caption{\label{fig-vRatios}Asymptotic velocity for each species, bulk velocity and normalized one by Alfvén velocity for the case of \(\lambda_{z,2} = - 1\) (a) and \(\lambda_{z,2} = - 1/2\) (b).}

\end{figure}%

\subsection{Current Sheet Thickness}\label{current-sheet-thickness}

TODO: define and compare current sheet thickness

TODO: compare with Harris current sheet solution with same thickness

\subsection{Density Peaking}\label{density-peaking}

\section{Summary}\label{summary}

\begin{figure}

\centering{

\includegraphics{article_files/mediabag/vl_ratio.png}

}

\caption{\label{fig-vRatioStats}Statistics of the asymptotic velocity ratio from PSP, Wind, and ARTEMIS spacecraft observations during PSP encounter 7 period from 2021-01-14 to 2021-01-21.}

\end{figure}%

  \bibliography{files/bibliography/research.bib}

\end{document}

### 2024-08-20 article.qmd
update articles

@@
 
-::: {.column-page}
-::: {#fig-examples layout-nrow="1" }
+::: {#fig-examples}
 
-Three examples of solar wind discontinuities observed by Parker Solar Probe (PSP), Wind and ARTEMIS spacecraft.
+![](figures/fig_examples.svg)
 
-:::
+Three examples of solar wind discontinuities observed by Parker Solar Probe (PSP), ARTEMIS and Wind spacecraft with sub-Alfvenic flow velocity.
+
 :::
@@
 
-::: {.column-page}
-::: {#fig-profiles layout-nrow="1" }
+::: {#fig-profiles layout-ncol="2" }
 
-![](figures/profiles_sym.svg)
+![](figures/profiles_sym.svg){#fig-profilesEx1}
 
-![](figures/J_profiles_sym.svg)
+![](figures/J_profiles_sym.svg){#fig-jprofilesEx1}
 
-(a). Magnetic field, ion density, and ion bulk velocity for $n_1 = n_2, u_1 = -u_2$ and $δ_1=1, c_1=1/\sqrt{2}, B_0 = 2$. (b). Current density profiles for the same case.
-:::
+(a). Magnetic field, ion density, and ion bulk velocity for $n_1 = n_2, u_1 = -u_2$ and $δ_1=1, c_1=1, B_0 = 2$. (b). Current density profiles for the same case.
+
 :::
@@
 
-We normalize the plasma velocity by asympotic Alfvén velocity $v_{A}(∞) = B_0 / \sqrt{\mu_0 m_i n}$, and the profiles are plotted below @fig-profilesC.
-It could be seen that for $n_1(∞) = 0.5$, we have zero bulk velocity change across the current sheet in the asymptotic limit.
-And the change of bulk velocity across the current sheet increases with $n_1(∞)$ decreasing to zero @fig-uxnorm.
+::: {#fig-profilesEx2 layout-ncol="2" }
+
+![](figures/profiles_n1Inf=0.6.svg)
+
+![](figures/J_profiles_n1Inf=0.6.svg)
+
+Same as @fig-profiles, but for two ion species with same bulk velocity but different densities $n_1(∞) = 0.6$ and $B_y(z=\infty) = 1/2 B_0$.
+
+:::
+
+We normalize the plasma velocity by asympotic Alfvén velocity $v_{A}(∞) = B_0 / \sqrt{\mu_0 m_i n}$, and the profiles are plotted in @fig-profilesEx2 for the case of $n_1(∞) = 0.6$. 
+@fig-uxnorm It could be seen that for $n_1(∞) = 0.5$, we have zero bulk velocity change across the current sheet in the asymptotic limit.
+And the change of bulk velocity across the current sheet increases with $n_1(∞)$ decreasing to zero.
 The normalized plasma velocity $U_y$ in the aympotic limit decreases with $n_1(∞)$ decreasing to zero @fig-uynorm.
 
-::: {.column-page}
 ::: {#fig-profilesC layout-nrow="1" }
-
-![Magnetic field, ion density, and ion bulk velocity for two ion species with same bulk velocity but different densities.](figures/profiles_ratio=0.5.svg){#fig-profiles_ratio}
 
@@
 :::
-:::
 
@@
 
-::: {.column-page}
+<!-- ::: {.column-page} -->
 ::: {#fig-uProfilesNorm layout-nrow="1" }
@@
 :::
-:::
+<!-- ::: -->
 
@@
 
-::: {.column-page}
+<!-- ::: {.column-page} -->
 ::: {#fig-vRatios layout-nrow="1" }

### 2024-08-27 article.qmd
Fix typo in article title

@@
 ---
-title: On a Alvenicity of solar wind discontinuities
+title: On a Alfvenicity of solar wind discontinuities
 authors:

### 2024-09-28 article.qmd
update: comparison

@@
     href: https://www.overleaf.com/project/668273f5d14067a22517615c
+  - text: Roam Research
+    href: https://roamresearch.com/#/offline/Research/page/5EiGYXsXE
+  - text: Zotero
+    href: zotero://select/library/collections/DZHNUFNV
 abstract: "Solar wind plasma flow is characterized by intense, large-amplitude current sheets, also known as solar wind discontinuities.
@@
 
-# References
-
-- @steinhauerMultifluidModelOnedimensional2008
-- @artemyevIonNongyrotropySolar2020
-- @neukirchKineticModelsTangential2020
-- @vaskoThinCurrentSheets2014
-
 <!-- "This difference has previously been interpreted as: (1) a possible contribution of anisotropy that decreases ΔvA, or (2) a discontinuity non-stationarity due to residual magnetic energy.
@@
   - Harris current sheet @harrisPlasmaSheathSeparating1962
+
+|      | Harris Current Sheet                              | Yoon-Lui Current Sheet                               | Kan Current Sheet                        |
+|-------------------|-------------------------------------------------------|--------------------------------------------------------|-----------------------------------------------------|
+| **Magnetic Field** | $B_x(z) = B_0 \tanh(z/L)$ | $B_x(z) = B_0 \tanh \left(\frac{h}{L} \tan^{-1}(z/h)\right)$|$B_x^*(x^*, z^*)$ |
+| **Number Density** | $n(z) = n_0 \text{sech}^2(z/L)$ | $n(z) = n_0 \text{sech}^2 \left(\frac{h}{L} \tan^{-1}(z/h)\right)$ | $n^*(x^*, z^*)$ |
+| **Current Density** | $j_y(z) = B_0 \text{sech}^2(z/L) / (\mu_0 L)$ |$j_y(z) = \frac{B_0}{\mu_0 L} \frac{h^2}{z^2 + h^2} \text{sech}^2 \left(\frac{h}{L} \tan^{-1}(z/h)\right)$|                     |
+| **Drift Velocity** |  |$V_y(z) = V_0 \frac{h^2}{(z^2 + h^2)}$ |                                       |
+| **Special Feature** | Vlasov equilibrium, Maxwellian velocity distribution | Velocity shear important for thin current sheets       | Function form involves complex variables            |
 

### 2024-09-30 article.qmd
Update multi-fluid model for force-free current sheet equilibria

@@
 
-|      | Harris Current Sheet                              | Yoon-Lui Current Sheet                               | Kan Current Sheet                        |
-|-------------------|-------------------------------------------------------|--------------------------------------------------------|-----------------------------------------------------|
-| **Magnetic Field** | $B_x(z) = B_0 \tanh(z/L)$ | $B_x(z) = B_0 \tanh \left(\frac{h}{L} \tan^{-1}(z/h)\right)$|$B_x^*(x^*, z^*)$ |
-| **Number Density** | $n(z) = n_0 \text{sech}^2(z/L)$ | $n(z) = n_0 \text{sech}^2 \left(\frac{h}{L} \tan^{-1}(z/h)\right)$ | $n^*(x^*, z^*)$ |
-| **Current Density** | $j_y(z) = B_0 \text{sech}^2(z/L) / (\mu_0 L)$ |$j_y(z) = \frac{B_0}{\mu_0 L} \frac{h^2}{z^2 + h^2} \text{sech}^2 \left(\frac{h}{L} \tan^{-1}(z/h)\right)$|                     |
-| **Drift Velocity** |  |$V_y(z) = V_0 \frac{h^2}{(z^2 + h^2)}$ |                                       |
-| **Special Feature** | Vlasov equilibrium, Maxwellian velocity distribution | Velocity shear important for thin current sheets       | Function form involves complex variables            |
-
 Observations of magnetic discontinuities exhibit Alfvenic properties, with the plasma velocity variation correlating well with magnetic field components variations.
@@
 ::: {#fig-examples}
-
 ![](figures/fig_examples.svg)
@@
 Three examples of solar wind discontinuities observed by Parker Solar Probe (PSP), ARTEMIS and Wind spacecraft with sub-Alfvenic flow velocity.
-
 :::
@@
 
-We used a steady-state multi-fluid model for assumed 1-D force-free current sheets.
+We used a multi-fluid model for assumed 1-D quasineutral force-free current sheet equilibria.
 This approach, inspired by @steinhauerMultifluidModelOnedimensional2008, has the advantage of being analytically tractable while providing the necessary complexity to include the kinetic effects of multiple ion species.
@@
 
+In this model, we assume that the asymptotic boundary conditions are known from measurements, including the normal component of the magnetic field and the asymptotic properties of the incident fluid streams, which are clearly distinguishable.
+It should be noted, however, that while the velocity space boundary between different groups may be clear in the uniform field region outside the current sheet, within the sheet itself, the boundary can become more complex.
+For instance, individual ions may jump from one group to another, and some may become trapped within the current sheet.
+These complexities, particularly ion behavior within the sheet, are beyond the scope of this paper and are not addressed in the current analysis.
+
 <!-- An example is the simultaneous appearance of separate beams of ions in different directions with different fluxes.
@@
 # Plasma-Field Equations for Force-free current sheet
+
+The governing equations for the multi-fluid plasma model can be expressed in conservation law form as follows:
+
+$$
+(\frac{∂}{∂ t} + \mathbf{u}_{α} · ∇) n_α = - n_α ∇ · \mathbf{u}_α
+$$
+
+$$
+m_α n_α (\frac{∂}{∂ t} + \mathbf{u}_{α} · ∇) \mathbf{u}_α  = - ∇ · \mathbf{P}_α + q_α n_α \left( \mathbf{E} + \frac{\mathbf{u}_α \times \mathbf{B}}{c} \right) + \mathbf{R}_α
+$$
+
+where $α$ indicates the particle species, $n_α$ the number density, $m_α$ the mass, $q_α$ the charge, $\mathbf{u}_α$ the bulk velocity, $\mathbf{P}_α$ the pressure tensor, $\mathbf{E}$ the electric field, $\mathbf{B}$ the magnetic induction (magnetic field), and $R_α$ the friction forces.
+
+In this study, we focus on the equilibrium solutions of the set of equations, which implies that the partial derivatives with respect to time are set to zero ($\frac{∂}{∂ t}=0$). 
+Our primary interest lies in the current sheet, and we simplify the problem by assuming that the magnetic field depends only on a single Cartesian coordinate, designated as $z$.
+Additionally, we assume that the background electron is massless, meaning the electrons follow the magnetic field lines $\mathbf{u}_e \times \mathbf{B} = 0$. 
+For further simplification, we also assume the ion populations are gyrotropic, eliminating any non-diagonal terms in the pressure tensor.
 
@@
 \end{aligned}
-$${#eq-momentum}
+$$ {#eq-momentum}
 
@@
 $$
-d B_y / d z = - J_x = - \sum_α n_α u_{α x} + n_e u_{e x} = -n u_x+\Gamma_e B_x / B_z
-$${#eq-Jx}
+d B_y / d z = - J_x = - \sum_α n_α u_{α x} + n_e u_{e x} = -n u_x + Γ_e B_x / B_z
+$$ {#eq-Jx}
 
 $$
-d B_x / d z = J_y = \sum_α n_α u_{α y} - n_e u_{e y} = n u_y-\Gamma_e B_y / B_z
-$${#eq-Jy}
+d B_x / d z = J_y = \sum_α n_α u_{α y} - n_e u_{e y} = n u_y - Γ_e B_y / B_z
+$$ {#eq-Jy}
 
 $$
-d B_z / d z = J_z = \sum_α n_α u_{α z} - n_e u_{e z} = n u_z-\Gamma_e
-$${#eq-Jz}
+d B_z / d z = J_z = \sum_α n_α u_{α z} - n_e u_{e z} = n u_z - Γ_e
+$$ {#eq-Jz}
 
@@
 n_α (u_{αx} B_y - n_α u_{αy} B_x) = 0
-$${#eq-alpha}
+$$ {#eq-alpha}
 
@@
 Conveniently, the center of the current sheet is chosen as the origin $z = 0$, which corresponds to the lower boundary of @eq-theta1.
-As a result, the boundary condition for the rotation angle is given by  $\theta_\alpha(0) = \pi/2$.
+As a result, the boundary condition for the rotation angle is given by  $\theta_α(0) = \pi/2$.
 
@@
 ::: {#fig-profiles layout-ncol="2" }
-
 ![](figures/profiles_sym.svg){#fig-profilesEx1}
@@
 (a). Magnetic field, ion density, and ion bulk velocity for $n_1 = n_2, u_1 = -u_2$ and $δ_1=1, c_1=1, B_0 = 2$. (b). Current density profiles for the same case.
-
 :::
@@
 ::: {#fig-profilesEx2 layout-ncol="2" }
-
 ![](figures/profiles_n1Inf=0.6.svg)
@@
 Same as @fig-profiles, but for two ion species with same bulk velocity but different densities $n_1(∞) = 0.6$ and $B_y(z=\infty) = 1/2 B_0$.
-
 :::
@@
 ::: {#fig-profilesC layout-nrow="1" }
-
 ![Plasma velocity $U_x$ profiles normalized by aysmpotic Alfvén velocity for different $n_1(∞)$.](figures/UxNormB0.svg){#fig-uxnorm}
@@
 ![Plasma velocity $U_y$ profiles normalized by aysmpotic Alfvén velocity for different $n_1(∞)$.](figures/UyNormB0.svg){#fig-uynorm}
-
 :::

### 2024-09-30 article.qmd
Update multi-fluid model for force-free current sheet equilibria

@@
 
-The governing equations for the multi-fluid plasma model can be expressed in conservation law form as follows:
+The governing equations for the multi-fluid collisionless plasma model can be expressed in conservation law form as follows:
 
@@
 $$
-m_α n_α (\frac{∂}{∂ t} + \mathbf{u}_{α} · ∇) \mathbf{u}_α  = - ∇ · \mathbf{P}_α + q_α n_α \left( \mathbf{E} + \frac{\mathbf{u}_α \times \mathbf{B}}{c} \right) + \mathbf{R}_α
+m_α n_α (\frac{∂}{∂ t} + \mathbf{u}_{α} · ∇) \mathbf{u}_α  = - ∇ · \mathbf{P}_α + q_α n_α \left( \mathbf{E} + \frac{\mathbf{u}_α \times \mathbf{B}}{c} \right)
 $$
 
-where $α$ indicates the particle species, $n_α$ the number density, $m_α$ the mass, $q_α$ the charge, $\mathbf{u}_α$ the bulk velocity, $\mathbf{P}_α$ the pressure tensor, $\mathbf{E}$ the electric field, $\mathbf{B}$ the magnetic induction (magnetic field), and $R_α$ the friction forces.
+where $α$ indicates the particle species, $n_α$ the number density, $m_α$ the mass, $q_α$ the charge, $\mathbf{u}_α$ the bulk velocity, $\mathbf{P}_α$ the pressure tensor, $\mathbf{E}$ the electric field a $\mathbf{B}$ the magnetic induction (magnetic field).
 
 In this study, we focus on the equilibrium solutions of the set of equations, which implies that the partial derivatives with respect to time are set to zero ($\frac{∂}{∂ t}=0$).
-Our primary interest lies in the current sheet, and we simplify the problem by assuming that the magnetic field depends only on a single Cartesian coordinate, designated as $z$.
+Since our primary interest lies in the force-free current sheet, we assume that the magnetic field depends only on a single Cartesian coordinate, $z$, leading to the force-free condition $B_x^2 + B_y^2 = B_0^2 = \text{const}$.
 Additionally, we assume that the background electron is massless, meaning the electrons follow the magnetic field lines $\mathbf{u}_e \times \mathbf{B} = 0$.
@@
 
-Following @steinhauerMultifluidModelOnedimensional2008, we consider a one-dimensional force-free current sheet where $B_x^2 + B_y^2 = B_0^2 = \text{const}$, with all variables dependent solely on $z$.
-The system comprises multiple steady-state ion species (fluid groups) and a background electron fluid.
-
-Electron motion is along the field lines $\vect{u_e} = Γ_e \vect{B} / (n_e B_z)$.
-And the conservation of fluid mass $d(n_α u_{αz})/dz = 0$ integrate to a constant parameters $Γ_α = n_α u_{αz}$.
-The momentum equations for each ion species are given by
+Under these assumptions, the conservation of fluid mass, expressed as $d(n_α u_{αz})/dz = 0$, integrates to a constant parameter, $Γ_α = n_α u_{αz}$.
+The electron motion is aligned with the magnetic field lines, resulting in $\vect{u_e} = Γ_e \vect{B} / (n_e B_z)$.
+The momentum equations for each ion species are then given by:
 

### 2024-10-02 article.qmd
Update dimension form

@@
 \newcommand{\vect}[1]{\mathbf{#1}}
+\DeclareMathOperator{\sech}{sech}
 :::
@@
 $$
-n_1(z) \to n_1 (∞) (\frac{c_1}{\frac{z}{\delta_1 }^2+1}+1)
+n_1(z) \to n_1(∞) +  \frac{κ Γ_1}{1+(z/L)^2}
 $$
@@
 $$
-\theta (z)\to \frac{\pi  \Gamma _1-2 c_1 \delta _1 B_z n_1(\infty ) \tan ^{-1}\left(\frac{z}{\delta _1}\right)}{2 \Gamma _1}
+\theta (z)\to \frac{\pi}{2} - \kappa  L B_z \tan ^{-1}\left(\frac{z}{L}\right)
 $$
 
-For the simplest situation where we have two group of ions of the same density but opposite bulk velocity, i.e., $n_1 = n_2, u_1 = -u_2$, the system could be determined given $c_1$, $δ_1$, and $B_0$.
+And therefore
+
+$$
+\begin{aligned}
+B_x(z) &= B_0 \sin \left(\kappa  L B_z \tan ^{-1}\left(\frac{z}{L}\right)\right)
+\\
+B_y(z) &= B_0 \cos \left(\kappa  L B_z \tan ^{-1}\left(\frac{z}{L}\right)\right)
+\\
+n(z) &= n(\infty ) + \frac{\kappa ∑_α Γ_α}{1+(z/L)^2}
+\\
+J_x(z) &= \frac{\kappa B_0 B_z}{1+(z/L)^2} \sin \left(\kappa  L B_z \tan ^{-1}\left(\frac{z}{L}\right)\right)
+\\
+J_y(z) &= \frac{\kappa B_0 B_z}{1+(z/L)^2} \cos \left(\kappa  L B_z \tan ^{-1}\left(\frac{z}{L}\right)\right)
+\end{aligned}
+$$
+
+For the simplest situation where we have two group of ions of the same density but opposite bulk velocity, i.e., $n_1 = n_2, u_1 = -u_2$, the system could be determined given $κ$, $L$, and $B_0$.
 The system profile of magnetic field, plasma density, and plasma velocity for one specific case ($δ_1=1, c_1=1/\sqrt{2}, B_0 = 2$) is plotted below @fig-profiles.
@@
 
+In non-dimensionalized form
+$$
+\begin{aligned}
+B_x(z) &= B_0 \sin \left(\frac{\kappa L}{d_i} \tan ^{-1}\left(\frac{z}{L}\right)\right)
+\\
+B_y(z) &= B_0 \cos \left(\frac{\kappa L}{d_i} \tan ^{-1}\left(\frac{z}{L}\right)\right)
+\\
+n(z) &= n(\infty ) + \frac{\kappa}{V_a} \frac{∑_α Γ_α}{1+(z/L)^2}
+\\
+J_x(z) &= \frac{\kappa B_0}{\mu_0 d_i}\frac{1}{1+(z/L)^2} \sin \left(\frac{\kappa L}{d_i} \tan ^{-1}\left(\frac{z}{L}\right)\right)
+\\
+J_y(z) &= \frac{\kappa B_0}{\mu_0 d_i}\frac{1}{1+(z/L)^2} \cos \left(\frac{\kappa L}{d_i} \tan ^{-1}\left(\frac{z}{L}\right)\right)
+\end{aligned}
+$$
+
 ::: {#fig-vRatioStats}

### 2024-10-04 article.qmd
update: SPARTHRB presentation

@@
 ---
-title: On a Alfvenicity of solar wind discontinuities
+title: On the Alfvenicity of solar wind discontinuities
 authors:
@@
 \newcommand{\vect}[1]{\mathbf{#1}}
-\DeclareMathOperator{\sech}{sech}
 :::
@@
 
-1. Observation features of solar wind discontinuities
-2. The necessity of kinetic models to describe solar wind discontinuities
-3. Current theoretical models of solar wind discontinuities and their limitations
+- Observation features of solar wind discontinuities
+- Importance of solar wind discontinuities
+- The necessity of kinetic models to describe solar wind discontinuities
+- Current theoretical models of solar wind discontinuities and their limitations
   - Harris current sheet @harrisPlasmaSheathSeparating1962
 
-Observations of magnetic discontinuities exhibit Alfvenic properties, with the plasma velocity variation correlating well with magnetic field components variations.
+<!-- Observations of magnetic discontinuities exhibit Alfvenic properties, with the plasma velocity variation correlating well with magnetic field components variations.
 However, the amplitude of the plasma velocity jump is often significantly smaller than the Alfvén velocity jump, indicating a more complicated picture than the classical MHD model of rotational discontinuities.
@@
 
-These differences could not be simply explained by the aniostropy of the plasma thermal pressure alone (assuming a Maxwellian distribution that fits both the parallel and perpendicular temperatures) and suggest that kinetic mechanisms are necessary to describe the stress balance within discontinuities.
+These differences could not be simply explained by the aniostropy of the plasma thermal pressure alone [@shenComparingPlasmaAnisotropy2024] (assuming a Maxwellian distribution that fits both the parallel and perpendicular temperatures) and suggest that kinetic mechanisms are necessary to describe the stress balance within discontinuities.
+-->
 
@@
 
-Three examples of solar wind discontinuities observed by Parker Solar Probe (PSP), ARTEMIS and Wind spacecraft with sub-Alfvenic flow velocity.
+Three examples of current sheets observed by Parker Solar Probe (PSP), ARTEMIS and Wind spacecraft with sub-Alfvenic flow velocity.
 :::
-
-Force-free @harrisonRemarksOnedimensionalForcefree2009 (A necessary condition on the pseudopotential plasma pressure to allow for force-free 1D VM equilibria is formulated.)
 
@@
 
+$$
+\mathbf {\nabla } \times \mathbf {B} =\mu _{0}\mathbf {J}
+$$ {#eq-Ampere}
+
+$$
+∇ · \mathbf{B} = 0
+$$
+
 where $α$ indicates the particle species, $n_α$ the number density, $m_α$ the mass, $q_α$ the charge, $\mathbf{u}_α$ the bulk velocity, $\mathbf{P}_α$ the pressure tensor, $\mathbf{E}$ the electric field a $\mathbf{B}$ the magnetic induction (magnetic field).
@@
 $$
-d B_z / d z = J_z = \sum_α n_α u_{α z} - n_e u_{e z} = n u_z - Γ_e
+0 = J_z = \sum_α n_α u_{α z} - n_e u_{e z} = n u_z - Γ_e
 $$ {#eq-Jz}
@@
 - B_0 \sin θ θ' = \sum n_α u_{α} \sin θ_α - Γ_e B_0 \sin θ / B_z
-% Λ_y n_1 u_1 \sin θ_1 - Λ_z Γ_1 B_0 \sin θ / B_z
 $$
@@
 For the simplest situation where we have two group of ions of the same density but opposite bulk velocity, i.e., $n_1 = n_2, u_1 = -u_2$, the system could be determined given $κ$, $L$, and $B_0$.
-The system profile of magnetic field, plasma density, and plasma velocity for one specific case ($δ_1=1, c_1=1/\sqrt{2}, B_0 = 2$) is plotted below @fig-profiles.
+The system profile of magnetic field, plasma density, and plasma velocity for one specific case ($L=1, κ=1, B_0 = 2$) is plotted below @fig-profiles.
 The specific profiles are chosen to have zero electron current across the current sheet and zero $B_y$ in the asymptotic limit, corresponding to a 180° rotation of the magnetic field across the current sheet.
@@
 
-(a). Magnetic field, ion density, and ion bulk velocity for $n_1 = n_2, u_1 = -u_2$ and $δ_1=1, c_1=1, B_0 = 2$. (b). Current density profiles for the same case.
+(a). Magnetic field, ion density, and ion bulk velocity for $n_1 = n_2, u_1 = -u_2$ and $L=1, κ=1, B_0 = 2$. (b). Current density profiles for the same case.
 :::
@@
 
-The asymptotic velocity for the each species, bulk velocity and normalized one by Alfvén velocity are plotted below @fig-vRatios and @fig-vRatiosEx2 for the case of $\lambda_{z,2} = - 1$ (two groups of opposite velocity in z direction) and $\lambda_{z,2} = - 1/2$ respectively.
-For the case of $\lambda_{z,2} = - 1$ as shown in @fig-vRatiosEx1, the normalized velocity reaches zero when $n_1(∞) = 0.5$.
+The asymptotic velocity for the each species, bulk velocity and normalized one by Alfvén velocity are plotted below @fig-vRatios and @fig-vRatiosEx2 for the case of $u_{1,z} = - u_{2,z}$ (two groups of opposite velocity in z direction) and $u_{1,z} = - 2 u_{2,z}$ respectively.
+For the case of $u_{1,z} = - u_{2,z}$ as shown in @fig-vRatiosEx1, the normalized velocity reaches zero when $n_1(∞) = 0.5$.
 The bulk velocity of each species is the same in the asymptotic limit and does not depend on $n_1(∞)$.
-However, this does not generally hold when $\lambda_{z,2} \neq - 1$ as shown in @fig-vRatiosEx2 for the case of $\lambda_{z,2} = - 1/2$..
+However, this does not generally hold when $u_{1,z} \neq - u_{2,z}$ as shown in @fig-vRatiosEx2 for the case of $u_{1,z} = - 2 u_{2,z}$.
 The normalized velocity reaches zero when $n_1(∞) = 1/3$, and the bulk velocity of each species differs in the asymptotic limit and depends on $n_1(∞)$.
@@
 
-Asymptotic velocity for each species, bulk velocity and normalized one by Alfvén velocity for the case of $\lambda_{z,2} = - 1$ (a) and $\lambda_{z,2} = - 1/2$ (b).
+Asymptotic velocity for each species, bulk velocity and normalized one by Alfvén velocity for the case of $u_{1,z} = - u_{2,z}$ (a) and $u_{1,z} = - 2 u_{2,z}$ (b).
 :::
@@
 \\
-n(z) &= n(\infty ) + \frac{\kappa}{V_a} \frac{∑_α Γ_α}{1+(z/L)^2}
+n(z) &= n_{\infty } + \frac{\kappa}{V_A} \frac{∑_α Γ_α}{1+(z/L)^2}
 \\

### 2024-10-04 index.qmd
update manuscripts and notebooks
update: SPARTHRB presentation

@@
 ---
-title: "Multifluid model of a one-dimensional steady state force-free current sheet"
+title: "Multifluid model of one-dimensional force-free current sheet equilibrium"
 format:
@@
   pptx:
-    slide-level: 3
+    slide-level: 2
     incremental: true
@@
 
-:::{.hide}
+::: hide
 \newcommand{\vect}[1]{\mathbf{#1}}
@@
 
-## Introduction
+# Introduction
 
-### Motivation
+Classification of discontinuities
 
-In MHD theory, for rotational rotational discontinuity, we have this following relationship
+The ratio $R_{V B}$ is expected to equal unity for rotational discontinuity, but is arbitrary for tangential discontinuities.
 
 $$
-[[\vec{U}]]=\left(\frac{\xi \rho}{\mu_0}\right)^{1 / 2}\left[\left[\frac{\vec{B}}{\rho}\right]\right]
+R_{V B} = \frac{1}{\sqrt{A}} \frac{|Δ \vect{V}|}{|Δ \vect{V_A}|}
 $$
 
-where $\xi \equiv 1-\frac{p_{\parallel}-p_{\perp}}{B^2 / \mu_0}$.
+where $Δ \vect{V_A} = Δ {\frac{\vect{B}}{\sqrt{\mu_0 \rho}}}$ and the anisotropy factor A is defined as $A = 1 - \frac{p_{\|}-p_{\perp}}{B^2/\mu_0}$.
+
+## Examples
+
+![Three examples of current sheets observed by Parker Solar Probe (PSP), ARTEMIS and Wind spacecraft with sub-Alfvenic flow velocity.](figures/fig_examples.svg)
 
@@
 
-However, observations of solar wind discontinuities reveals discrepancies between
+## Force-free current sheets
 
-- Alfven velocity and plasma velocity change across discontinuities
+Force-free current sheets are important for modelling low-$β$ plasmas in laboratory, space and astrophysical applications [@marshForcefreeMagneticFields1996]
+
+However, observations of current sheet reveals discrepancies between
+
+- Alfven velocity and plasma velocity change across current sheet
 - Anisotropic MHD theory-predicted and directly measured ion anisotropies
 
-Multifluid theory is proposed to address these discrepancies. The basic idea is that we could have zero bulk velocity with non zero pressure.
+## Theory - previous art
 
-#### MHD in a nutshell
+### Multi-fluid model
+
+A multi-fluid model is developed to estimate the ratio of $R_{V B}$ [@neugebauerReexaminationRotationalTangential1984]. However no analytical solution is provided for the force-free current sheet.
 
 $$
-\begin{aligned}
-& \frac{\partial \rho}{\partial t}+(\vec{V} \cdot \nabla) \rho+\rho \nabla \cdot \vec{V}=0 \\
-& \rho\left[\frac{\partial \vec{V}}{\partial t}+(\vec{V} \cdot \nabla) \vec{V}\right]+\nabla p=\vec{J} \times \vec{B} \\
-& p=\alpha \rho^\gamma \\
-& \nabla \times \vec{B}=\mu_0 \vec{J} \\
-& \frac{\partial \vec{B}}{\partial t}=\nabla \times(\vec{V} \times \vec{B})
-\end{aligned}
+R_{V B}(\text { multistream })=\left(\frac{1}{A} \sum_i \frac{|Δ \vect{V_i}|^2}{|Δ \vect{V_{i,A}} |^2}\right)^{1 / 2}
 $$
-
-### Previous Work
-
-On the extreme of multiple fluid theory, we have the kinetic Harris model for the current sheet.
-
-- no normal field, $B_z = 0$
-- uniform cross-tail drift velocity, $u_y$
-- vanishing particle density far from the sheet
-
-<!-- > Consider several deficiencies of the Harris model: (1) there is no normal field, Bz = 0, so that no particles cross the sheet from asymptotic sources; (2) the unrealistic assumption of uniform cross-tail drift velocity, uy; (3) vanishing particle density far from the sheet; (4) assumption of a simple Maxwellian velocity distribution; (5) further, one study showed that atypical current sheet structures with bifurcated current density occur 17% of the time, and even for typical center-peaked sheets, strongly peaked structures frequently appear in the center, which also differs from the Harris profile [Asano et al., 2005].
--->
-
-@steinhauerMultifluidModelOnedimensional2008 built a multifluid model to study the steady state magnetotail current sheet.
 
@@
 
-For the symmetric three-fluid system with $B_y=0$,
+### Steinhauer, McCarthy, and Whipple (2008)
+
+@steinhauerMultifluidModelOnedimensional2008 built a multifluid model to study the steady state magnetotail current sheet.
+
+For the symmetric three-fluid system with $B_y=0$. However, it is not force-free.
 
@@
 
-Here, we are interested in the case of a force-free current sheet and start from the simplest case of two ion species.
+---
 
-## Plasma-Field Equations for a One-Dimensional Multifluid
+### Vlasov equilibrium
 
-Consider a collisionless, steady state plasma composed of multiple fluid groups, and follow the notation in @steinhauerMultifluidModelOnedimensional2008, we have $3N$ equations for $N$ ion species:
+A number of Vlasov-Maxwell distribution function have been found to describe force-free current sheet recently
+
+$$
+\begin{aligned}
+& \frac{\mathrm{d}^2 A_x}{\mathrm{d} z^2}=-\mu_0 \frac{\partial P_{z z}}{\partial A_x} \\
+& \frac{\mathrm{d}^2 A_y}{\mathrm{d} z^2}=-\mu_0 \frac{\partial P_{z z}}{\partial A_y}
+\end{aligned}
+$$
+
+::: {.notes}
+As has been shown before, the 1D VM equilibrium equations are equivalent to the motion of a pseudoparticle in a conservative pseudopotential, with the pseudopotential being proportional to one of the diagonal components of the plasma pressure tensor.
+:::
+
+- no normal field, $B_z = 0$
+
+<!-- > Consider several deficiencies of the Harris model: (1) there is no normal field, Bz = 0, so that no particles cross the sheet from asymptotic sources; (2) the unrealistic assumption of uniform cross-tail drift velocity, uy; (3) vanishing particle density far from the sheet; (4) assumption of a simple Maxwellian velocity distribution; (5) further, one study showed that atypical current sheet structures with bifurcated current density occur 17% of the time, and even for typical center-peaked sheets, strongly peaked structures frequently appear in the center, which also differs from the Harris profile [Asano et al., 2005].
+-->
+
+# Plasma-Field Equations
+
+## Multi-fluid collisionless plasma model in a nutshell
+
+The governing equations for the multi-fluid collisionless plasma model can be expressed in conservation law form as follows:
+
+$$
+(\frac{∂}{∂ t} + \mathbf{u}_{α} · ∇) n_α = - n_α ∇ · \mathbf{u}_α
+$$
+
+$$
+m_α n_α (\frac{∂}{∂ t} + \mathbf{u}_{α} · ∇) \mathbf{u}_α  = - ∇ · \mathbf{P}_α + q_α n_α \left( \mathbf{E} + \frac{\mathbf{u}_α \times \mathbf{B}}{c} \right)
+$$
+
+$$
+\mathbf {\nabla } \times \mathbf {B} =\mu _{0}\mathbf {J}
+$$ {#eq-Ampere}
+
+$$
+∇ · \mathbf{B} = 0
+$$
+
+where $α$ indicates the particle species...
+
+---
+
+## To much, let's simplify
+
+- equilibrium solutions => $\frac{∂}{∂ t}=0$
+
+- current sheet => magnetic field depends only on a single Cartesian coordinate, $z$
+
+- force-free condition => $B_x^2 + B_y^2 = B_0^2 = \text{const}$
+
+Additionally, we assume
+
+- background electron is massless => electrons follow the magnetic field lines $\mathbf{u}_e \times \mathbf{B} = 0$
+
+- ion populations are gyrotropic => eliminating any non-diagonal terms in the pressure tensor
+
+To further simplify, we employ a dimensionless system by normalizing the variables with their asymptotic values:
+
+- magnetic field by $B_{\text{ref}} = B_z$
+
+- density by $n_{\text{ref}} = n(∞)$
+
+---
+
+## After all the simplifications
+
+Conservation of fluid mass integrates to a constant parameter $Γ_α = n_α u_{αz}$;
+
+Gauss's law gives another constant parameter $B_{z} = \text{const}$;
+
+The momentum equations for each ion species are then given by:
 
@@
 $$
-d B_y / d z = - J_x = - \sum_α n_α u_{α x} + n_e u_{e x} = -n u_x+\Gamma_e B_x / B_z
-$${#eq-Jx}
+d B_y / d z = - J_x = - \sum_α n_α u_{α x} + n_e u_{e x} = -n u_x + Γ_e B_x / B_z
+$$ {#eq-Jx}
 
 $$
-d B_x / d z = J_y = \sum_α n_α u_{α y} - n_e u_{e y} = n u_y-\Gamma_e B_y / B_z
-$${#eq-Jy}
+d B_x / d z = J_y = \sum_α n_α u_{α y} - n_e u_{e y} = n u_y - Γ_e B_y / B_z
+$$ {#eq-Jy}
 
@@
 
+<!-- 
 ::: {#def-1}
@@
 - $Λ_z$ is constant while $Λ_x, Λ_y$ are not.
-
 ::: 
+-->
 
@@
 
-$C_1=0$ immediately implies $θ = θ_1 + k \pi$ and $Λ_x=Λ_y$. #TODO: Prove $C_1=0$ is like a boundary condition not a equation.
-
 The first momentum equation (@eq-momentum) after substituting the above expression becomes:
@@
 \Rightarrow 
-θ_α' = - \frac{n_α B_z}{Γ_α} \pm \frac{B_0}{u_α}
+θ_α' = - \frac{n_α B_z}{Γ_α} + \frac{B_0}{u_α}
 $${#eq-theta1}
@@
 - **So given the profile of $n_α$, we could solve the above equation to get the profile of $θ_α$, thus the profile of $u_α$.**
-- The derivative of $θ$ goes to zero at infinity, gives us a relation between $u_1$ and $B_0$:
+- The derivative of $θ$ goes to zero at infinity, gives us a relation between $u_α$ and $B_0$:
 
   $$
-  u_1 = \pm \frac{B_0 Γ_1}{B_z n_1(\infty)}
+  u_α = \pm \frac{B_0 Γ_α}{B_z n_α(\infty)}
   $${#eq-u1}
@@
 $$
-- B_0 \sin θ θ' = Λ_y n_1 u_1 \sin θ_1 - Λ_z Γ_1 B_0 \sin θ / B_z
+- B_0 \sin θ θ' = \sum n_α u_{α} \sin θ_α - Γ_e B_0 \sin θ / B_z
 $$
@@
 \Rightarrow 
-θ' = \mp \frac{Λ_y n_1 u_1 }{B_0} + \frac{Λ_z Γ_1}{B_z}
+θ' = - \frac{ \sum n_α u_{α} }{B_0} + \frac{Γ_e}{B_z}
 $${#eq-theta}
 
-By equating the above two equations, we could get a relation between $n_1$ and $Λ_y$:
+The above equation relates the rotation angle $θ$ to the plasma bulk velocity $n u \equiv \sum n_α u_{α}$.
+By equating the above two equations of $θ'$, we could get a equation relating the plasma bulk velocity to one specific species:
 
 $$
-Λ_y = Λ_y(n_1) = -\frac{B_0 \left(-n_1 u_1 B_z^2+B_0 \Gamma _1 B_z-\Gamma _1^2 u_1 \Lambda _z\right)}{\Gamma _1 n_1 u_1^2 B_z}
+n u = B_0 \left(\frac{Γ_e}{B_z}+\frac{n_α B_z}{Γ_α}-\frac{B_0}{u_α}\right)
 $$
 
-### Solutions
-
-Normalize the density by $n_1(z=\infty) = 1$ and the magnetic field by $B_z = 1$.
+## Solutions
 
@@
 
-Assuming density profile $n_1(z) \to \frac{c}{\left(\frac{z}{\delta }\right)^2+1}+1$
+Assuming density profile
 
-We have
+$$
+n_1(z) \to n_1(∞) +  \frac{κ Γ_1}{1+(z/L)^2}
+$$
+
+Solving the @eq-theta1-asym analytically, we have the rotation angle profile as
+
+$$
+\theta (z)\to \frac{\pi}{2} - \kappa  L B_z \tan ^{-1}\left(\frac{z}{L}\right)
+$$
+
+---
+
+## Beutiful moment - solutions
+
+:::: {.columns}
+
+::: column
+In dimensionless form
 
@@
 \begin{aligned}
-\varphi (z) &\to \frac{\pi  \Gamma _1-2 c \delta  \tan ^{-1}\left(\frac{z}{\delta }\right)}{2 \Gamma _1}
+& B_x(z) = B_0 \sin \left(\kappa  L B_z \tan ^{-1}\left(\frac{z}{L}\right)\right)
 \\
-B_x &\to B_0 \cos \left(\frac{\pi  \Gamma _1-2 c \delta  \tan ^{-1}\left(\frac{z}{\delta }\right)}{2 \Gamma _1}\right)
+& B_y(z) = B_0 \cos \left(\kappa  L B_z \tan ^{-1}\left(\frac{z}{L}\right)\right)
+\\
+& n(z) = n(\infty ) + \frac{\kappa ∑_α Γ_α}{1+(z/L)^2}
+\\
+& J_x(z) = \frac{\kappa B_0 B_z}{1+(z/L)^2} \sin \left(\kappa  L B_z \tan ^{-1}\left(\frac{z}{L}\right)\right)
+\\
+& J_y(z) = \frac{\kappa B_0 B_z}{1+(z/L)^2} \cos \left(\kappa  L B_z \tan ^{-1}\left(\frac{z}{L}\right)\right)
 \end{aligned}
 $$
+:::
+
+::: column
+In physical units
+
+$$
+\begin{aligned}
+& B_x(z) = B_0 \sin \left(\frac{\kappa L}{d_i} \tan ^{-1}\left(\frac{z}{L}\right)\right)
+\\
+& B_y(z) = B_0 \cos \left(\frac{\kappa L}{d_i} \tan ^{-1}\left(\frac{z}{L}\right)\right)
+\\
+& n(z) = n_{\infty } + \frac{\kappa}{V_A} \frac{∑_α Γ_α}{1+(z/L)^2}
+\\
+& J_x(z) = \frac{\kappa B_0}{\mu_0 d_i}\frac{1}{1+(z/L)^2} \sin \left(\frac{\kappa L}{d_i} \tan ^{-1}\left(\frac{z}{L}\right)\right)
+\\
+& J_y(z) = \frac{\kappa B_0}{\mu_0 d_i}\frac{1}{1+(z/L)^2} \cos \left(\frac{\kappa L}{d_i} \tan ^{-1}\left(\frac{z}{L}\right)\right)
+\end{aligned}
+$$
+:::
+::::
+
+---
 
@@
 
-![](figures/J_profiles.jpg)
+![](figures/J_profiles_sym.svg)
 
 The multi-fluid effect is clearly demonstrated by $Λ_y - 1$.
-
-```mathematica
-Manipulate[
- Block[{B0 = B0i, c = ci},
-  p1 = Plot[
-    Evaluate[{n[z], Bx[z], By[z], u1x[z], u1y[z]} /. sols //. rulesn],
-    {z, -zmax, zmax},
-    PlotLegends -> {n, Subscript[B, x], Subscript[B, y], Subscript[u, x], Subscript[u, y]}
-  ];
-  Export["figures/profiles.jpg", p1];
-  p1
-  ],
-
- {{ci, 1/Sqrt[2]}, 0, 2},
- {{B0i, 2}, 0.5, 2},
- {{zmax, 10}, 5, 30}
- ]
-
-Manipulate[
- Block[{B0 = B0i, c = ci, Λz = Λz0},
-  p = Plot[
-    Evaluate[{Λy - 1, Jxi, Jx, Jyi, Jy} /. sols //. 
-      rulesn],
-    {z, -zmax, zmax},
-    PlotLegends -> {Subscript[Λ, y] - 1, Subscript[J, 
-      x, i], Subscript[J, x], Subscript[J, y, i], Subscript[J, y]}, 
-    PlotRange -> All];
-  Export["figures/J_profiles.jpg", p];
-  p
-  ],
-
- {{ci, 1/Sqrt[2]}, 0, 2},
- {{B0i, 2}, 0.5, 2},
- {Λz0, 0, 1},
- {{zmax, 10}, 5, 30}
-]
-```
 
@@
 
-::: {.column-screen-inset}
-::: {#fig-profiles layout-nrow="1" }
+---
 
-![Magnetic field, ion density, and out ion bulk velocity for $λ_{n,∞}=1$ and $λ_{z,∞}=-1$.](figures/profiles_ratio=0.5.svg){#fig-profiles_ratio}
+## Symmetric case
 
-![Normalized Plasma Velocity $U_x$ Profiles for Different $λ_{n,∞}$.](figures/UxNorm.svg){#fig-uxnorm}
+![](figures/profiles_sym.svg){#fig-profilesEx1}
 
-![Normalized Plasma Velocity $U_y$ Profiles for Different $λ_{n,∞}$.](figures/UyNorm.svg){#fig-uynorm}
+![](figures/J_profiles_sym.svg){#fig-jprofilesEx1}
 
-:::
-:::
+(a). Magnetic field, ion density, and ion bulk velocity for $n_1 = n_2, u_1 = -u_2$ and $δ_1=1, c_1=1, B_0 = 2$. (b). Current density profiles for the same case.
 
-![](figures/vRatios.svg)
+## Asymmetric case
 
-## Different normalization
+![](figures/profiles_n1Inf=0.6.svg)
 
-Normalized the quantities of the ion species
+![](figures/J_profiles_n1Inf=0.6.svg)
 
-$$
-n_α := λ_{αn} n_e
-\\
-u_{αx} := λ_{αu} B_z / \sqrt{n_{e,\infty}}
-\\
-u_{αy} := λ_{αu} B_z / \sqrt{n_{e,\infty}}
-\\
-u_{αz} := λ_{αu} B_z / \sqrt{n_{e,\infty}}
-$$
+Same as @fig-profiles, but for two ion species with same bulk velocity but different densities $n_1(∞) = 0.6$ and $B_y(z=\infty) = 1/2 B_0$.
 
-$$
-θ' = \mp \frac{Λ_u}{B_0} + Γ_e
-$$
+---
 
-```mathematica
-conf = {ratio -> 1/2}
-TextString[conf]
-<!-- Convert a dict-like configuration to a string -->
-saveName[c_] := StringJoin[ToString /@ c]
-```
+![Plasma velocity $U_x$ profiles normalized by aysmpotic Alfvén velocity for different $n_1(∞)$.](figures/UxNormB0.svg){#fig-uxnorm}
+
+![Plasma velocity $U_y$ profiles normalized by aysmpotic Alfvén velocity for different $n_1(∞)$.](figures/UyNormB0.svg){#fig-uynorm}
+
+---
+
+![Plasma velocity $U_x$ profiles normalized by $v_{A,x}$ for different $n_1(∞)$.](figures/UxNormBx.svg){#fig-uxNormBx}
+
+![Plasma velocity $U_y$ profiles normalized by $v_{A,y}$ for different $n_1(∞)$.](figures/UyNormBy.svg){#fig-uyNormBy}
+
+---
+
+![](figures/vRatios.svg){#fig-vRatiosEx1}
+
+![](figures/vRatiosEx2.svg){#fig-vRatiosEx2}
+
+Asymptotic velocity for each species, bulk velocity and normalized one by Alfvén velocity for the case of $u_{1,z} = - u_{2,z}$ (a) and $u_{1,z} = - 2 u_{2,z}$ (b).

### 2024-10-04 article.qmd
update

--- 2024-09-30 article.qmd
@@
 
-In this model, we assume that the asymptotic boundary conditions are known from measurements, including the normal component of the magnetic field and the asymptotic properties of the incident fluid streams, which are clearly distinguishable.
+In this model, we assume that the asymptotic boundary conditions are known from measurements, including the normal component of the magnetic field, the magnitude of the magnetic field, and the asymptotic properties of the incident fluid streams.
 It should be noted, however, that while the velocity space boundary between different groups may be clear in the uniform field region outside the current sheet, within the sheet itself, the boundary can become more complex.
@@
 
-The multifluid model developed above is used here to study a specific example and the effects of model parameters: assume a density profile for one species taking the form of a Gaussian function
+The multifluid model developed above is used here to study a specific example and the effects of model parameters: assume a density profile for one species taking the form of a Lorentzian function with a background density.
 

### 2024-11-18 presentations/AGU24.qmd
feat: add poster

--- 2024-07-22 index.qmd
@@
 ---
-title: "Multifluid model of a one-dimensional steady state force-free current sheet"
-format:
-  html: default
-  pptx:
-    slide-level: 3
-    incremental: true
-code-links:
-  - text: Tana
-    href: https://app.tana.inc?nodeid=OzzX0sVA4_6X
-  - text: Zotero
-    href: zotero://select/library/collections/DZHNUFNV
+title: Multifluid model of one-dimensional force-free current sheet equilibrium
+subtitle: AGU24 poster
+author:
+  - name: Zijin Zhang
+    affil: 1
+  - name: Anton Artemyev
+    affil: 1
+  - name: Marco Velli
+    affil: 1
+  - name: Vassilis Angelopoulos
+    affil: 1
+affiliation:
+  - num: 1
+    address: Department of Earth, Planetary, and Space Sciences, University of California, Los Angeles
+output: "posterdown::posterdown_betterland"
 ---
 
-:::{.hide}
-\newcommand{\vect}[1]{\mathbf{#1}}
-:::
+## Introduction & Motivation
 
-## Introduction
+'Discontinuities' are discontinuous spatial changes in plasma
+parameters/characteristics and magnetic fields.
 
-### Motivation
+## Method
 
-In MHD theory, for rotational rotational discontinuity, we have this following relationship
-
-$$
-[[\vec{U}]]=\left(\frac{\xi \rho}{\mu_0}\right)^{1 / 2}\left[\left[\frac{\vec{B}}{\rho}\right]\right]
-$$
-
-where $\xi \equiv 1-\frac{p_{\parallel}-p_{\perp}}{B^2 / \mu_0}$.
-
----
-
-However, observations of solar wind discontinuities reveals discrepancies between
-
-- Alfven velocity and plasma velocity change across discontinuities
-- Anisotropic MHD theory-predicted and directly measured ion anisotropies
-
-Multifluid theory is proposed to address these discrepancies. The basic idea is that we could have zero bulk velocity with non zero pressure.
-
-#### MHD in a nutshell
+## Results
 
@@
 \begin{aligned}
-& \frac{\partial \rho}{\partial t}+(\vec{V} \cdot \nabla) \rho+\rho \nabla \cdot \vec{V}=0 \\
-& \rho\left[\frac{\partial \vec{V}}{\partial t}+(\vec{V} \cdot \nabla) \vec{V}\right]+\nabla p=\vec{J} \times \vec{B} \\
-& p=\alpha \rho^\gamma \\
-& \nabla \times \vec{B}=\mu_0 \vec{J} \\
-& \frac{\partial \vec{B}}{\partial t}=\nabla \times(\vec{V} \times \vec{B})
+& B_x(z) = B_0 \sin \left(\frac{\kappa L}{d_i} \tan ^{-1}\left(\frac{z}{L}\right)\right)
+\\
+& B_y(z) = B_0 \cos \left(\frac{\kappa L}{d_i} \tan ^{-1}\left(\frac{z}{L}\right)\right)
+\\
+& n(z) = n_{\infty } + \frac{\kappa}{V_A} \frac{∑_α Γ_α}{1+(z/L)^2}
+\\
+& J_x(z) = \frac{\kappa B_0}{\mu_0 d_i}\frac{1}{1+(z/L)^2} \sin \left(\frac{\kappa L}{d_i} \tan ^{-1}\left(\frac{z}{L}\right)\right)
+\\
+& J_y(z) = \frac{\kappa B_0}{\mu_0 d_i}\frac{1}{1+(z/L)^2} \cos \left(\frac{\kappa L}{d_i} \tan ^{-1}\left(\frac{z}{L}\right)\right)
 \end{aligned}
@@
 
-### Previous Work
+::: {#fig-profiles layout-ncol="2" }
+![](../figures/profiles_sym.svg){#fig-profilesEx1}
 
-On the extreme of multiple fluid theory, we have the kinetic Harris model for the current sheet.
+![](../figures/J_profiles_sym.svg){#fig-jprofilesEx1}
 
-- no normal field, $B_z = 0$
-- uniform cross-tail drift velocity, $u_y$
-- vanishing particle density far from the sheet
-
-<!-- > Consider several deficiencies of the Harris model: (1) there is no normal field, Bz = 0, so that no particles cross the sheet from asymptotic sources; (2) the unrealistic assumption of uniform cross-tail drift velocity, uy; (3) vanishing particle density far from the sheet; (4) assumption of a simple Maxwellian velocity distribution; (5) further, one study showed that atypical current sheet structures with bifurcated current density occur 17% of the time, and even for typical center-peaked sheets, strongly peaked structures frequently appear in the center, which also differs from the Harris profile [Asano et al., 2005].
--->
-
-@steinhauerMultifluidModelOnedimensional2008 built a multifluid model to study the steady state magnetotail current sheet.
-
-![Ion trajectories in model magnetic field.](figures/jgra18999-fig-0001.png)
-
----
-
-For the symmetric three-fluid system with $B_y=0$,
-
-![Magnetic field, current density, ion density and out-of-plane flow for Bz = 0.1, βe = 0.5, βi = 1.
-Solid lines are fluid model results and dot symbols are from the Harris sheet solution.](figures/jgra18999-fig-0003.png)
-
-Here, we are interested in the case of a force-free current sheet and start from the simplest case of two ion species.
-
-## Plasma-Field Equations for a One-Dimensional Multifluid
-
-Consider a collisionless, steady state plasma composed of multiple fluid groups, and follow the notation in @steinhauerMultifluidModelOnedimensional2008, we have $3N$ equations for $N$ ion species:
-
-$$
-\begin{aligned}
-Γ_α \frac{d u_{α x}}{d z} &= n_α u_{α y} B_z - Γ_α B_y
-\\
-Γ_α \frac{d u_{α y}}{d z} &= Γ_α B_x-n_α u_{α x} B_z
-\\
-Γ_α \frac{d u_{α z}}{d z} &= -\frac{1}{2} \frac{d p_α}{d z}-n_α \frac{d \phi}{d z}+n_α u_{α x} B_y-n_α u_{α y} B_x
-\end{aligned}
-$${#eq-momentum}
-
-Ampere’s law connects the fields and the flow components and note that electron motion is along the field lines $\vect{u_e} = Γ_e \vect{B} / (n_e B_z)$ :
-
-$$
-d B_y / d z = - J_x = - \sum_α n_α u_{α x} + n_e u_{e x} = -n u_x+\Gamma_e B_x / B_z
-$${#eq-Jx}
-
-$$
-d B_x / d z = J_y = \sum_α n_α u_{α y} - n_e u_{e y} = n u_y-\Gamma_e B_y / B_z
-$${#eq-Jy}
-
----
-
-This is a system of $3N+2$ equations for $4N+2$ unknown dependent variables: $B_x$, $B_y$, and $N$ each of $n$, $p$ $u_x$, and $u_y$.
-
-- $B_z, Γ_α$ are constant.
-
-- Combine momentum equation (x) (@eq-momentum) and Ampere's law (y) (@eq-Jy) with condition that the constant of integration vanishes at the current sheet center yield
-
-  $$
-  B_x B_z = \sum n_α u_{α,x} u_{α,z} = \sum Γ_α u_{α,x}
-  $$
-
-  - With asymptotic condition that the derivative goes zero, from momentum equation we have
-
-    $$
-    {B_z}^2 = \sum \Gamma_{\alpha}^2/n_{\alpha}(\infty)
-    $${#eq-asym1}
-
-In the following, we will consider a system with two ion species.
-
-## Force-free current sheet
-
-We are interested in solutions with $B_x^2 + B_y^2 = B_0^2 = const$ (force-free current sheet).
-This provides another equation for the system, now we have 3*2+2+1=9 equations for 10 unknowns (the system is close to be fully determined).
-From Ampere's law (@eq-Jx) and (@eq-Jy), we have:
-
-$$
-n (u_x B_y  - u_y B_x) = 0
-$$
-
-And $n_1 u_{1x} B_y = n_1 u_{1y} B_x + C_1$ and $n_2 u_{2x} B_y = n_2 u_{2y} B_x  - C_1$. Assuming $C_1 = 0$, from first two equations in (@eq-momentum), we have:
-
-$$
-u_{αx}^2 + u_{αy}^2 = const
-$$
-
----
-
-::: {#def-1}
-Express the quantities of the second ion species relative to the first species
-
-$$
-u_{αx} := λ_{αx} u_{1x}, \quad u_{αy} := λ_{αy} u_{1y}, \quad u_{αz} := λ_{αz} u_{1z}, \quad n_α := λ_{αn} n_1
-$$
-
-$$
-Λ_x := 1 + \sum λ_{αx} λ_{αn}, \quad Λ_y := 1 + \sum λ_{αy} λ_{αn}, \quad Λ_z := \sum 1 + λ_{αz} λ_{αn}
-$$
-
-Notes
-
-- $Γ_e = (1 + λ_z λ_n) Λ_1 =  Λ_z  Γ_1$. 
-- $Λ_z$ is constant while $Λ_x, Λ_y$ are not.
-
+(a). Magnetic field, ion density, and ion bulk velocity for $n_1 = n_2, u_1 = -u_2$ and $L=1, κ=1, B_0 = 2$. (b). Current density profiles for the same case.
 :::
 
-::: {#def-2}
-Force-free condition let us express the magnetic field and velocity in terms of the angle $θ$:
+::: {#fig-profilesEx2 layout-ncol="2" }
+![](../figures/profiles_n1Inf=0.6.svg)
 
-$$
-B_x = B_0 \cos θ, \quad B_y = B_0 \sin θ
-$$
+![](../figures/J_profiles_n1Inf=0.6.svg)
 
-$$
-u_{1x} = u_{1} \cos θ_1, \quad u_{1y} = u_{1} \sin θ_1
-$$
+Same as @fig-profiles, but for two ion species with same bulk velocity but different densities $n_1(∞) = 0.6$ and $B_y(z=\infty) = 1/2 B_0$.
 :::
 
----
+::: {#fig-vRatios layout-nrow="1" }
+![](../figures/vRatios.svg){#fig-vRatiosEx1}
 
-$C_1=0$ immediately implies $θ = θ_1 + k \pi$ and $Λ_x=Λ_y$. #TODO: Prove $C_1=0$ is like a boundary condition not a equation.
+![](../figures/vRatiosEx2.svg){#fig-vRatiosEx2}
 
-The first momentum equation (@eq-momentum) after substituting the above expression becomes:
-
-$$
-- u_α \sin θ_α θ_α' = n_α u_α \sin θ_α B_z / Γ_α - B_0 \sin θ
-$$
-
-$$
-\Rightarrow 
-θ_α' = - \frac{n_α B_z}{Γ_α} \pm \frac{B_0}{u_α}
-$${#eq-theta1}
-
-- Note that $θ_α, n_α$ are dependent variables, and $u_α, B_0, B_z, Γ_α$ are constants determined by the system.
-- **So given the profile of $n_α$, we could solve the above equation to get the profile of $θ_α$, thus the profile of $u_α$.**
-- The derivative of $θ$ goes to zero at infinity, gives us a relation between $u_1$ and $B_0$:
-
-  $$
-  u_1 = \pm \frac{B_0 Γ_1}{B_z n_1(\infty)}
-  $${#eq-u1}
-
-The Ampere's law (@eq-Jy) become:
-
-$$
-- B_0 \sin θ θ' = Λ_y n_1 u_1 \sin θ_1 - Λ_z Γ_1 B_0 \sin θ / B_z
-$$
-
-$$
-\Rightarrow 
-θ' = \mp \frac{Λ_y n_1 u_1 }{B_0} + \frac{Λ_z Γ_1}{B_z}
-$${#eq-theta}
-
-By equating the above two equations, we could get a relation between $n_1$ and $Λ_y$:
-
-$$
-Λ_y = Λ_y(n_1) = -\frac{B_0 \left(-n_1 u_1 B_z^2+B_0 \Gamma _1 B_z-\Gamma _1^2 u_1 \Lambda _z\right)}{\Gamma _1 n_1 u_1^2 B_z}
-$$
-
-### Solutions
-
-Normalize the density by $n_1(z=\infty) = 1$ and the magnetic field by $B_z = 1$.
-
-The system $S$ could be fully determined by $λ_n, λ_z, B_0$ parameters, provided $n(z)$ profiles.
-
-$$
-S=S(z; n(z); λ_n, λ_z, B_0)
-$$
-
-Assuming density profile $n_1(z) \to \frac{c}{\left(\frac{z}{\delta }\right)^2+1}+1$
-
-We have
-
-$$
-\begin{aligned}
-\varphi (z) &\to \frac{\pi  \Gamma _1-2 c \delta  \tan ^{-1}\left(\frac{z}{\delta }\right)}{2 \Gamma _1}
-\\
-B_x &\to B_0 \cos \left(\frac{\pi  \Gamma _1-2 c \delta  \tan ^{-1}\left(\frac{z}{\delta }\right)}{2 \Gamma _1}\right)
-\end{aligned}
-$$
-
-For the simplest case $λ_n = 1, λ_z = - 1$, from @eq-asym1, we have $Γ_1 = - B_z / \sqrt{2}$.
-
-We could also normalize the system length by $\delta = 1$. So now the system could be fully determined by $c, B_0$.
-
-Given $c = 1/\sqrt{2}$, we have $B_y \to 0$ as $z \to \infty$. Profiles are plotted below for $B_0 = 2$.
-
-![](figures/profiles.jpg)
-
-![](figures/J_profiles.jpg)
-
-The multi-fluid effect is clearly demonstrated by $Λ_y - 1$.
-
-```mathematica
-Manipulate[
- Block[{B0 = B0i, c = ci},
-  p1 = Plot[
-    Evaluate[{n[z], Bx[z], By[z], u1x[z], u1y[z]} /. sols //. rulesn],
-    {z, -zmax, zmax},
-    PlotLegends -> {n, Subscript[B, x], Subscript[B, y], Subscript[u, x], Subscript[u, y]}
-  ];
-  Export["figures/profiles.jpg", p1];
-  p1
-  ],
-
- {{ci, 1/Sqrt[2]}, 0, 2},
- {{B0i, 2}, 0.5, 2},
- {{zmax, 10}, 5, 30}
- ]
-
-Manipulate[
- Block[{B0 = B0i, c = ci, Λz = Λz0},
-  p = Plot[
-    Evaluate[{Λy - 1, Jxi, Jx, Jyi, Jy} /. sols //. 
-      rulesn],
-    {z, -zmax, zmax},
-    PlotLegends -> {Subscript[Λ, y] - 1, Subscript[J, 
-      x, i], Subscript[J, x], Subscript[J, y, i], Subscript[J, y]}, 
-    PlotRange -> All];
-  Export["figures/J_profiles.jpg", p];
-  p
-  ],
-
- {{ci, 1/Sqrt[2]}, 0, 2},
- {{B0i, 2}, 0.5, 2},
- {Λz0, 0, 1},
- {{zmax, 10}, 5, 30}
-]
-```
-
-For the same asymptotic magnetic field, it is interesting to see how the plasma profiles change with different system parameters.
-
-$$
-\text{cond}(n(z), λ_{n,∞}, λ_{z,∞}) = 0
-$$
-
-Here we set $B_y(z=\infty) = 1/2 B_y(z=0)$ and $B_0 = 2 B_z$, and fix $λ_{z,∞} = -1$.
-By varying $λ_{n,∞}$, we find that the magnetic field profiles are exactly the same, while plasma velocity profiles vary.
-We normalize the plasma velocity by asympotic Alfvén velocity $v_{A,∞} = B_0 / \sqrt{n}$, and the profiles are plotted below @fig-uxnorm.
-It could be seen that for $λ_{n,∞} = 1$, we have zero bulk velocity change across the current sheet in the asymptotic limit.
-
-::: {.column-screen-inset}
-::: {#fig-profiles layout-nrow="1" }
-
-![Magnetic field, ion density, and out ion bulk velocity for $λ_{n,∞}=1$ and $λ_{z,∞}=-1$.](figures/profiles_ratio=0.5.svg){#fig-profiles_ratio}
-
-![Normalized Plasma Velocity $U_x$ Profiles for Different $λ_{n,∞}$.](figures/UxNorm.svg){#fig-uxnorm}
-
-![Normalized Plasma Velocity $U_y$ Profiles for Different $λ_{n,∞}$.](figures/UyNorm.svg){#fig-uynorm}
-
+Asymptotic velocity for each species, bulk velocity and normalized one by Alfvén velocity for the case of $u_{1,z} = - u_{2,z}$ (a) and $u_{1,z} = - 2 u_{2,z}$ (b).
 :::
-:::
-
-![](figures/vRatios.svg)
-
-## Different normalization
-
-Normalized the quantities of the ion species
-
-$$
-n_α := λ_{αn} n_e
-\\
-u_{αx} := λ_{αu} B_z / \sqrt{n_{e,\infty}}
-\\
-u_{αy} := λ_{αu} B_z / \sqrt{n_{e,\infty}}
-\\
-u_{αz} := λ_{αu} B_z / \sqrt{n_{e,\infty}}
-$$
-
-$$
-θ' = \mp \frac{Λ_u}{B_0} + Γ_e
-$$
-
-```mathematica
-conf = {ratio -> 1/2}
-TextString[conf]
-<!-- Convert a dict-like configuration to a string -->
-saveName[c_] := StringJoin[ToString /@ c]
-```

### 2024-11-26 presentations/AGU24.qmd
feat: :art: AGU24 poster

@@
 output: "posterdown::posterdown_betterland"
+main_textsize: 120px
+main_findings:
+  - We developed a multi-fluid model for 1-D quasineutral force-free current sheet equilibria
+  - Our model explains the sub-Alfvénic velocity jumps observed in solar wind discontinuities through the interaction of interpenetrating ion beams
 ---
@@
 
-'Discontinuities' are discontinuous spatial changes in plasma
-parameters/characteristics and magnetic fields.
+Solar wind discontinuities are characterized by intense, large-amplitude current sheets that exhibit complex kinetic properties beyond classical MHD models.
+While these discontinuities show Alfvénic properties with correlated plasma and magnetic field variations, spacecraft observations reveal that plasma velocity jumps are often significantly smaller than Alfvén velocity jumps.
+This discrepancy cannot be explained by plasma pressure anisotropy alone, suggesting the need for kinetic mechanisms in describing the stress balance within discontinuities.
+
+![Three examples of current sheets observed by Parker Solar Probe (PSP), ARTEMIS and Wind spacecraft with sub-Alfvenic flow velocity.](../figures/fig_examples.svg)
 
@@
 
+We developed a multi-fluid model that:
+
+- Assumes 1-D quasineutral force-free current sheet equilibria
+
+- Incorporates multiple ion species with different bulk velocities
+
+- Uses asymptotic boundary conditions from observations
+
 ## Results
+
+Our model yields analytical solutions for the magnetic field, current density, and plasma parameters:
 
@@
 
-::: {#fig-profiles layout-ncol="2" }
-![](../figures/profiles_sym.svg){#fig-profilesEx1}
+where $B_0$: asymptotic field magnitude, $n_{\infty}$: asymptotic density, $V_A$: Alfvén speed, $\kappa$: density enhancement factor, $L$: current sheet width, $d_i = c/ω_{pi}$: ion inertial length, $Γ_α$: species mass flux
 
-![](../figures/J_profiles_sym.svg){#fig-jprofilesEx1}
+Asymmetric case: Profiles for two ion species with same bulk velocity but different densities ($n_1(∞) = 0.6$) and $B_y(z=\infty) = 1/2 B_0$.
 
-(a). Magnetic field, ion density, and ion bulk velocity for $n_1 = n_2, u_1 = -u_2$ and $L=1, κ=1, B_0 = 2$. (b). Current density profiles for the same case.
-:::
-
-::: {#fig-profilesEx2 layout-ncol="2" }
-![](../figures/profiles_n1Inf=0.6.svg)
-
-![](../figures/J_profiles_n1Inf=0.6.svg)
-
-Same as @fig-profiles, but for two ion species with same bulk velocity but different densities $n_1(∞) = 0.6$ and $B_y(z=\infty) = 1/2 B_0$.
-:::
-
-::: {#fig-vRatios layout-nrow="1" }
-![](../figures/vRatios.svg){#fig-vRatiosEx1}
-
-![](../figures/vRatiosEx2.svg){#fig-vRatiosEx2}
-
-Asymptotic velocity for each species, bulk velocity and normalized one by Alfvén velocity for the case of $u_{1,z} = - u_{2,z}$ (a) and $u_{1,z} = - 2 u_{2,z}$ (b).
-:::
+Velocity ratios analysis: (a) Symmetric case with $u_{1,z} = - u_{2,z}$ showing zero normalized velocity at $n_1(∞) = 0.5$.
+(b) Asymmetric case with $u_{1,z} = - 2u_{2,z}$ showing zero normalized velocity at $n_1(∞) = 1/3$, demonstrating how ion population asymmetries affect the bulk flow characteristics.

### 2024-12-06 article.qmd
feat: update fig-vRatios
update

@@
 
-The multifluid model developed above is used here to study a specific example and the effects of model parameters: assume a density profile for one species taking the form of a Lorentzian function with a background density.
+The multifluid model developed above is used here to study a specific example and the effects of model parameters.
+We assume a density profile for the first ion species ($n_1$) that takes the form of a Lorentzian function with a background density:
 
@@
 $$
+
+where $n_1(∞)$ is the asymptotic (background) density far from the current sheet, $κ$ is a dimensionless parameter controlling the peak density enhancement, and $L$ is the characteristic width of the current sheet.
 
@@
 
-<!-- ::: {.column-page} -->
 ::: {#fig-vRatios layout-nrow="1" }
-![](figures/vRatios.svg){#fig-vRatiosEx1}
+![](figures/vRatios.pdf){#fig-vRatiosEx1}
 
-![](figures/vRatiosEx2.svg){#fig-vRatiosEx2}
+![](figures/vRatiosEx2.pdf){#fig-vRatiosEx2}
 
@@
 :::
-<!-- ::: -->
 

### 2025-01-12 article.qmd
docs: update

@@
 
-- Observation features of solar wind discontinuities
+<!-- - Observation features of solar wind discontinuities
 - Importance of solar wind discontinuities
@@
 - Current theoretical models of solar wind discontinuities and their limitations
-  - Harris current sheet @harrisPlasmaSheathSeparating1962
+  - Harris current sheet @harrisPlasmaSheathSeparating1962 -->
 
@@
 
-Ampere’s law connects the fields and the flow components
+Ampere's law connects the fields and the flow components
 
 $$
-d B_y / d z = - J_x = - \sum_α n_α u_{α x} + n_e u_{e x} = -n u_x + Γ_e B_x / B_z
+\frac{1}{μ_0} \frac{d B_y}{d z}  = - J_x = - \sum_α n_α u_{α x} + n_e u_{e x} = -n U_x + Γ_e B_x / B_z
 $$ {#eq-Jx}
@@
 $$
-d B_x / d z = J_y = \sum_α n_α u_{α y} - n_e u_{e y} = n u_y - Γ_e B_y / B_z
+\frac{1}{μ_0} \frac{d B_x}{d z} = J_y = \sum_α q_α n_α u_{α y} - e n_e u_{e y} = e (n U_y - Γ_e \frac{B_y}{B_z})
 $$ {#eq-Jy}
@@
 $$
-0 = J_z = \sum_α n_α u_{α z} - n_e u_{e z} = n u_z - Γ_e
+0 = J_z = \sum_α n_α u_{α z} - n_e u_{e z} = n U_z - Γ_e
 $$ {#eq-Jz}
@@
 $$
-- u_α \sin θ_α θ_α' = n_α u_α \sin θ_α B_z / Γ_α - B_0 \sin θ
+- u_α \sin θ_α θ_α' = \frac{e}{m_α} (n_α u_α \sin θ_α B_z / Γ_α - B_0 \sin θ)
 $$
@@
 \Rightarrow 
-θ_α' = - \frac{n_α B_z}{Γ_α} + \frac{B_0}{u_α}
+θ' = \frac{e}{m_α} \left(\frac{B_0}{u_α} - \frac{n_α B_z}{Γ_α}\right)
 $${#eq-theta1}
+where $'=d/dz$.
 
-Note that $θ_α, n_α$ are dependent variables, and $u_α, B_0, B_z, Γ_α$ are constants pre-determined by the system.
-So given the profile of $n_α$ and the boundary condition, we could solve the above equation to get the profile of $θ_α$, thus the profile of $u_α$.
+Note that $θ, n_α$ are dependent variables, and $u_α, B_0, B_z, Γ_α$ are constants pre-determined by the system.
+So for a given profile of $n_α$ and the boundary condition, we could solve the above equation to get the profile of $θ$.
 
@@
 $$
-- B_0 \sin θ θ' = \sum n_α u_{α} \sin θ_α - Γ_e B_0 \sin θ / B_z
+- B_0 \sin θ θ' = μ_0 e \left(\sum n_α u_{α} \sin θ_α - Γ_e  \sin θ \frac{B_0}{B_z}\right)
 $$
@@
 \Rightarrow 
-θ' = - \frac{ \sum n_α u_{α} }{B_0} + \frac{Γ_e}{B_z}
+θ' = μ_0 e \left(\frac{Γ_e}{B_z} - \frac{ \sum n_α u_{α} }{B_0}\right)
 $${#eq-theta}
@@
 $$
-n u = B_0 \left(\frac{Γ_e}{B_z}+\frac{n_α B_z}{Γ_α}-\frac{B_0}{u_α}\right)
+n U = B_0 \left(\frac{Γ_e}{B_z}+\frac{1}{μ_0 m_α}\left(\frac{n_α B_z}{Γ_α}-\frac{B_0}{u_α}\right)\right)
 $$
@@
 In the asymptotic region, as all variables approach constant values, the derivatives must vanish.
-For @eq-theta1, this means $θ_α(∞)' = 0 = - \frac{n_α(∞) B_z}{Γ_α} + \frac{B_0}{u_α}$.
+For @eq-theta1, this means $0 = - \frac{n_α(∞) B_z}{Γ_α} + \frac{B_0}{u_α}$.
 This relates the velocity of each species to the asymptotic density.
@@
 $$
+where $\sum_α Γ_α=\sum_α n_α(∞)u_{z,α}(∞)$ and $i$ represents either $x$ or $y$.
+Using the relation that $Δn_α / Δn = Γ_α / \sum_α Γ_α$ and Equation~\ref{eq-asym1}, $\sum_α Γ_α$ can be reformulated into a form that does not involve $u_{z,α}(∞)$, which is often challenging to measure.
+This alternative expression is given by:
+
+$$
+\sum_α Γ_α = \sqrt{ \frac{B_z^2}{μ_0 m_p} \frac{{Δn}^2}{\sum_α {Δn_α}^2/ n_α(∞)} }
+$$
+
+For the simplest situation where we have two populations of ions of the same density but with oppositely directed bulk velocity, i.e., \(n_1 = n_2, u_1 = -u_2\), the system could be determined given $\frac{\sum_α Γ_α}{Δn}$, \(L\), and \(B_0\).
+The system profile of magnetic field, plasma density, and plasma velocity for one specific case ($L=d_i, \sum_α Γ_α= Δn V_A, B_0 = 2 B_z$) is plotted below Figure~\ref{fig-profiles}.
+The specific profiles are chosen to have zero electron current across the current sheet and zero \(B_y\) in the asymptotic limit (far away from the current sheet), corresponding to a $180^\circ$ rotation of the magnetic field across the current sheet.
+However, in general, we would expect the electron current to be non-zero.
 

### 2025-01-18 nbs/examples.py
chore: use script instead of notebooks

@@ -1,0 +1,93 @@
+# %% [markdown]
+# ## Get data
+%load_ext autoreload
+%autoreload 2
+from discontinuitypy.plot.alfvenicity import tsplot_Alfvenicity
+from space_analysis.ds.spz.io import to_dataarray
+import speasy as spz
+from speasy.core.requests_scheduling.request_dispatch import init_cdaweb
+import importlib.util
+import matplotlib.pyplot as plt
+from beforerr.matplotlib import PlotOpts, process_figure, easy_save
+import holoviews as hv
+
+# %%
+hv.extension("matplotlib")
+
+if importlib.util.find_spec("scienceplots") is not None:
+    importlib.import_module("scienceplots")
+    plt.style.use(["science", "nature", "notebook"])
+init_cdaweb()
+
+# %%
+plot_opts = PlotOpts(sync_legend_colors=True, hide_legend_lines=True)
+
+# %%
+coord = "gsm"
+thm_products = [
+    "cda/THB_L2_FGM/thb_fgs_gsm",
+    "cda/THB_L2_MOM/thb_peim_velocity_gsmQ",
+    "cda/THB_L2_MOM/thb_peim_densityQ",
+]
+thm_tr = ["2016-05-20T17:47:00", "2016-05-20T17:53:40"]
+
+psp_products = [
+    "cda/PSP_FLD_L2_MAG_RTN/psp_fld_l2_mag_RTN",
+    "cda/PSP_SWP_SPI_SF00_L3_MOM/VEL_SC",
+    "cda/PSP_SWP_SPI_SF00_L3_MOM/DENS",
+]
+
+# higher resolution but in `tplot` format
+psp_products = [
+    "psp_fld_l2_mag_RTN_4_Sa_per_Cyc",
+    "psp_swp_spi_af00_L3_VEL_RTN",
+    "psp_swp_spi_af00_L3_DENS",
+]
+
+psp_tr = ["2021-01-17T13:54:17", "2021-01-17T13:54:32"]
+
+wi_products = ["cda/WI_H2_MFI/BGSE", "cda/WI_PM_3DP/P_VELS", "cda/WI_PM_3DP/P_DENS"]
+wi_tr = ["2011-08-26T21:26:00", "2011-08-26T21:27:30"]
+
+# %%
+from psp.io.psp import get_data as psp_get_data
+
+
+def get_data(products, timerange):
+    for product in products:
+        if product in psp_products:
+            yield psp_get_data(product, timerange)
+        else:
[…]
